---
title: "Infections-Frailty-CPRD"
author: "Kwabena Asare"
date: "`r lubridate::today()`"
format: html
execute: 
  include: true
  echo: false
  eval: false
editor_options: 
  chunk_output_type: console
---

# Main Analyses

**Individuals making at least one follow-up**: this is the main analysis sample

## Descriptive results: Table one

**Table 1.** Descriptive summaries of **baseline** characteristics by severe infection (exposure) status.
**Table 2.** Descriptive summaries of **follow-up** eFI by severe infection (exposure) status.

```{r table_one_main}

# baseline characteristics by exposure status

table_one_main <- inf_frailty_cprd |>
  filter(if_any(matches("^eFI_yr[1-5]$"),~!is.na(.x))) |> 
  transmute(exposed = if_else(exposed == 0, "No","Yes"),
            age_bl,age_bl_cat4,gender,ethnicity,
            e2019_imd,harmful_alcohol_use,
            smoking,care_home_residency,
            gp_infection_5yr_lookback,gp_infection_7dy_lookback,
            antibio_7dy_lookbck,dementia,diabetes,
            atrial_fibrillation_bl,
            chronic_kidney_disease_bl,hypertension_bl,
            urinary_system_disease_bl,
            osteoporosis_bl,parkinsonism_tremors_bl,
            ischaemic_heart_disease_bl,heart_failure_bl,
            cerebrovascular_disease_bl,rheumatoid_diseases_bl,
            eFI_bl,eFI_bl_cat,one_consult_1y_lookbck) |> collect() |>
  rename_with(.fn = ~str_replace(.x,"_bl","")) |>
  tbl_summary(by="exposed",
              missing_text = "Missing",
              type=list(c(age,eFI) ~ "continuous2"),
              statistic = list(all_continuous2() ~ c("{mean} ({sd})","{median} ({p25}, {p75})")),
              digits = list(all_continuous() ~ 0,c(eFI) ~ 2)) |>
  bold_labels() |>
  modify_footnote(everything() ~ NA) |>
  modify_caption(glue("**Table 1.** Characteristics of participants at cohort entry (England)")) |>
  modify_spanning_header(all_stat_cols() ~ "**Severe infection after age 65 years**")

table_one_main
```

```{r table_two_main}

# follow-up eFI by exposure status
table_two_main <- inf_frailty_cprd |>
  filter(if_any(matches("^eFI_yr[1-5]$"),~!is.na(.x))) |> 
  select(person_time_years,
         exposed,eFI_bl,
         follow_up_yr1,eFI_yr1,
         follow_up_yr2,eFI_yr2,
         follow_up_yr3,eFI_yr3,
         follow_up_yr4,eFI_yr4,
         follow_up_yr5,eFI_yr5) |> 
  collect() |>
  mutate(exposed = if_else(exposed == 0, "No","Yes")) |>
  filter(!is.na(eFI_bl),
         if_any(matches("^eFI_yr[1-5]$"),~!is.na(.x))) |>
  select(!eFI_bl) |>
  tbl_summary(by="exposed",
              missing_text = "Missing", # Note: no deficit is missing; deficit = 1,0 if not censored
              type=list(c(eFI_yr1,eFI_yr2,eFI_yr3,eFI_yr4,eFI_yr5) ~ "continuous2"),
              statistic = list(all_continuous2() ~ c("{mean} ({sd})")),
              digits = list(all_continuous2() ~ 2,person_time_years~1),
              label = c(eFI_yr1 ~ "eFI (yr 1)",
                        eFI_yr2 ~ "eFI (yr 2)",
                        eFI_yr3 ~ "eFI (yr 3)",
                        eFI_yr4 ~ "eFI (yr 4)",
                        eFI_yr5 ~ "eFI (yr 5)",
                        follow_up_yr1 ~ "Follow-up (yr 1)",
                        follow_up_yr2 ~ "Follow-up (yr 2)",
                        follow_up_yr3 ~ "Follow-up (yr 3)",
                        follow_up_yr4 ~ "Follow-up (yr 4)",
                        follow_up_yr5 ~ "Follow-up (yr 5)",
                        person_time_years ~"Follow-up time, years")) |>
  bold_labels() |>
  modify_footnote(everything() ~ NA) |>
  modify_caption(glue("**Table 2.** eFI during follow-up (England)")) |>
  modify_spanning_header(all_stat_cols() ~ "**Severe infection after age 65 years at cohort entry**")

table_two_main
```

## Linear regression models (GEE): conditional mean difference in **Electronic Frailty Index (eFI)** at each of follow-up years 1, 2, 3, 4, and 5.
Using GEE for robust standard error estimation by match id = setid

### Main models

For each outcome build models with the following sequence of confounder adjustment sets:

 - **Minimally adjusted** models: adjust for matching variables (age, gender), ethnicity, and baseline eFI
 - **Deprivation adjusted** models: further adjust minimal model for:
    - individual and practice level IMD (quntiles)
 - **Lifestyle adjusted** models: further adjust deprivation model for:
    - smoking
    - harmful alcohol use
 - **Underlying health** models: further adjust lifestyle model for:
  - 5 yr look back infections; gp recorded

### Linear GEE regression function

Create one function to fit the following models:

- main cross-sectional models
- stratified EM models
- interaction models

**Function arguments**:
**outcome_var**, `chr` = outcome variable.
**cov_set**, `chr` = covariate set.
**remove_var**, `chr` = to specify any variable that needs to be removed from an already specified covariate set.
**interaction_var**, `chr` = if not NULL, fit include interaction term with main exposure. Note: remove exposure and interaction_var as individual covariates.
**data**= `inf_frailty_cprd` is the default. For EM stratified models, to supply stratified data sets.

```{r gee_model_function,eval=TRUE}
#
geelm <- function(outcome_var,
                  cov_set = c('matching','deprivation','lifestyle','underlying_health','ethnicity'),
                  remove_var=NULL,
                  interaction_var=NULL,
                  data=inf_frailty_cprd) { 

  # assert that remove_var, interaction_var is either NULL or a character vector
  stopifnot(is.null(remove_var) | is.character(remove_var))
  stopifnot(is.null(interaction_var) | is.character(interaction_var))

  # deFIne covariate sets 
  # switch covariate set based on cov_set specification
  # numeric all covariates for efficiency; use factor(as.numeric(.x)); if becomes necessary
  # check data dictionary for variable labels
  
  matching_eFI_bl <- c("exposed","age_bl_cat4","gender","eFI_bl") 
  deprivation <- c("e2019_imd")
  lifestyle <- c("harmful_alcohol_use","smoking")
  underlying_health <- c("gp_infection_5yr_lookback")
  ethnicity <- c("ethnicity") 

  
  # sequential confounder adjustment with covariate sets
  
  cov_set <- match.arg(cov_set) # this makes the first option the default and will throw an error if none is selected
  
  covariates <- switch(cov_set,
  "matching" = matching_eFI_bl, 
  "deprivation" = c(matching_eFI_bl,deprivation),          
  "lifestyle" = c(matching_eFI_bl,deprivation,lifestyle),         
  "underlying_health" = c(matching_eFI_bl,deprivation,lifestyle,underlying_health),
  "ethnicity" = c(matching_eFI_bl,deprivation,lifestyle,underlying_health,ethnicity),
  stop("select a cov_set: one of c('matching','deprivation','lifestyle','underlying_health',ethnicity)"))
  
  # if remove_var is supplied (i.e, not null) remove from covariate set
  # else, leave as is 

  if(!is.null(remove_var) && !remove_var %in% covariates) stop("Must select remove_var variable from the covariate set")
  
  covariates <- 
    if(!is.null(remove_var)) covariates[-c(match(remove_var,covariates))]  else {covariates} 
      
  # model data
  # tidy data based on whether running cross-section or longitudinal/repeated measures model 

  data <- collect(data) |> # manually remove missing observations: geeglm() does not automatically do that
  filter(if_all(c(all_of(outcome_var),all_of(covariates)),~(!is.na(.x) & str_detect(.x,"[Mm]issing",negate = T)))) |>
  select(c(setid,all_of(outcome_var),all_of(covariates),any_of(interaction_var))) 
  
  # model formula
  # if interaction is false then normal model
  # otherwise include interaction term and manually remove main exposure and interaction_var as individual covariates from model as the function geeglm() automatically adds
  # exclude dropped covariates
  
  covariates <- covariates[covariates %in% names(data)]

  if(!is.null(interaction_var) && !interaction_var %in% names(data)) stop("Must choose interaction_var from the dataset")

  f <- if(is.null(interaction_var))  {
    as.formula(paste(outcome_var,"~",c(glue_collapse(covariates,sep = "+"))))
    } else 
      
  if(!is.null(interaction_var)) {
    as.formula(paste0(outcome_var,"~","exposed*",interaction_var,"+",c(glue_collapse(
      covariates[-Filter(function(x) !is.na(x),match(c("exposed",interaction_var),covariates))],
                             sep = "+"))))
    }

  # Linear GEE model
  # for cross-section model use gee(independence corstr) for robust standard errors (id = setid)
  geeglm(f,
         corstr = "independence",
         family = gaussian(link = "identity"),
         id=setid,
         data = data)

  }

```

### Regression models

For each covariate set, fit yearly models and bundle model objects into a named list

```{r model_vars}
outcome_variables <- c("eFI_yr1","eFI_yr2","eFI_yr3","eFI_yr4","eFI_yr5")
model_years <- c("Year 1","Year 2","Year 3","Year 4","Year 5")
```

```{r matching_models}

matching_models <- setNames(map(outcome_variables,~geelm(outcome = .x,cov_set="matching"),.progress = T),
                           model_years)
```

```{r deprivation_models}
deprivation_models <- setNames(map(outcome_variables,~geelm(outcome = .x,cov_set="deprivation"),.progress = T),
                           model_years)
```

```{r lifestyle_models}
lifestyle_models <- setNames(map(outcome_variables,~geelm(outcome = .x,cov_set="lifestyle"),.progress = T),
                           model_years)
```

```{r underlying_health_models}
underlying_health_models <- setNames(map(outcome_variables,~geelm(outcome = .x,cov_set="underlying_health"),.progress = T),
                           model_years)
```

### Forest plots

```{r forest_plots}

# data.frame one: extract model elements into a dataframe
  df_models <- reduce(map(underlying_health_models,~tidy(.x,conf.int = T)[2,c("estimate", "conf.low","conf.high")]),
                      rbind) |>
    mutate(est_label = paste0(formatC(estimate,format = "f",digits = 3),
                              " (",formatC(conf.low,format = "f",digits = 3),", ",
                              formatC(conf.high,format = "f",digits = 3),")"),
           estimate = round(estimate,3)) |>
    mutate("Time" = factor(names(underlying_health_models)),.before = 1) |>
    mutate(est_label_middot=paste0(str_replace_all(est_label,"[.]","\U00B7")),
           estimate_middot = str_replace(estimate,"[.]","\U00B7")) |>
    add_row(Time = "Index date",.before = 1)

# data.frame two: percentage censored
  df_censor_death <- data.frame(Time = factor(c(sort(as.character(rep(df_models$Time,2))))),
                            Exposure = rep(c("No Severe infection at index date","Severe infection at index date"),6),
                            Censor_death = c(NA,NA,NA,NA,4.6,22,8.5,29,12,34,15,37)) |>
    mutate(Censor_death_middot=str_replace(Censor_death,"[.]","\U00B7"))
  
# data.frame three: descriptive mean eFI
  df_desc_means <- data.frame(Time = factor(c(sort(as.character(rep(df_models$Time,2))))),
                              Exposure = rep(c("No Severe infection at index date","Severe infection at index date"),6),
                              Mean = c(0.12,0.17,0.13,0.18,0.13,0.19,0.14,0.20,0.15,0.20,0.15,0.21)) |>
    mutate(Mean_middot=str_replace(formatC(Mean,format = "f",2),"[.]","\U00B7"))
    
  
# Follow-up plot: censor death, censor from other causes are similar in both exposure groups
  
# descriptive plus annotation plot
  fp_censor_death <- ggplot(df_censor_death,aes(Censor_death,Time,fill = factor(Exposure,levels = c("Severe infection at index date","No Severe infection at index date")))) + 
    geom_col(position = "dodge",width=0.4) + 
    geom_text(aes(x =Censor_death+3,label = if_else(!is.na(Censor_death),paste0(Censor_death,"%"),NA_character_)),
              position = position_dodge(width = 0.5),size = 5,fontface = "bold") +
    scale_x_continuous("Died (%)",breaks = seq(0,40,10),limits = c(0,45)) +
    scale_y_discrete(name=NULL,limits = rev) +
    scale_fill_discrete(guide = guide_legend("",direction = "horizontal",ncol=2)) +
    theme(text = element_text(face="bold",size=14),
          panel.border = element_rect(fill = NA),
          panel.grid.minor = element_blank(),
          legend.background = element_blank(),
          legend.position = "bottom",
          axis.ticks = element_line(linewidth=2),
          legend.text = element_text(size=14,face="bold"),
           legend.title = element_text(size=20,face="bold"),
          legend.key.size = unit(0.3, 'cm'),
          axis.text= element_text(size=14,face="bold",hjust = -0.1,colour = "black"),
          plot.margin = margin(20,15,10,30)) +
    labs(title = "",subtitle = "") # t,r,b,l
  
  # descriptive plus annotation plot
  fp_mean <- ggplot(df_desc_means,aes(Mean,Time,fill = factor(Exposure,levels = c("Severe infection at index date","No Severe infection at index date")))) + 
    geom_col(position = "dodge",width=0.4) + 
    geom_text(aes(x =Mean+0.01,label = Mean),size = 5,fontface = "bold",
              position = position_dodge(width = 0.5)) +
    scale_x_continuous("Mean eFI",limits = c(0,0.22)) +
    scale_y_discrete("",limits = rev,label=NULL) +
    scale_fill_discrete(guide = guide_legend("",direction = "horizontal",ncol=2)) +
    theme(text = element_text(face="bold",size=14),
          axis.text= element_text(size=14,face="bold",hjust = -0.1,colour = "black"),
          panel.border = element_rect(fill = NA),
          panel.grid.minor = element_blank(),
          legend.position = "bottom",
          legend.background = element_blank(),
          axis.ticks = element_line(linewidth=2),
          legend.text = element_text(size=14,face="bold"),
          legend.title = element_text(size=20,face="bold"),
          legend.key.size = unit(0.3, 'cm'),
          plot.margin = margin(20,15,10,15)) +
    labs(title = "Panel 2: English cohort",subtitle = "")

      
# regression plus annotation plot
  fp_regression <- ggplot(df_models,aes(estimate,Time)) + 
    geom_point(shape=18,size=7,colour="red") +
    geom_linerange(aes(xmin=conf.low,xmax=conf.high),linewidth=1) +
    geom_vline(xintercept = 0,linetype="dashed",colour="blue",linewidth=2) +
    geom_text(aes(x=estimate+0.02,label=est_label),
              fontface= "bold",size = 7) +
    scale_y_discrete(name=NULL,label=NULL,limits=rev) + 
    scale_x_continuous("Conditional mean difference in eFI (95% CI)",limits = c(0,0.05)) +
    theme(text = element_text(face="bold",size=14),
          axis.text= element_text(size=14,face="bold",hjust = -0.1,colour = "black"),
          panel.border = element_rect(fill = NA),
          panel.grid.minor = element_blank(),
          axis.ticks = element_line(linewidth=2),
          plot.margin = margin(20,10,10,30)) +
    labs(title = "",subtitle = "")

  # combine plots
  Figure_2 <- ggarrange(fp_censor_death,fp_mean,fp_regression,
                        allign="h",nrow=1,
                        labels=c("              A.","B.","C."),
                        label.y = 0.93,
                        common.legend = T,
                        legend = "none") 
  
```


### Table three: regression models

```{r merge_stack_model_tables_by_year,eval=TRUE}

# Create a function to merge model-tables by year: put yearly models for each covariate set into one row, in order of time

merge_tables_by_year <- function(models,model_name,
                                 label = "Sequential confounder adjustment",
                                 model_years = c("Year 1", "Year 2", "Year 3", "Year 4", "Year 5")){
  stopifnot(is.list(models)) # assert models are supplied as a list
  
  #
  tbl_merge( # to stack/append the tables
    map(models, 
        ~tbl_regression(.x,include="exposed",
                        label = list("exposed" ~ model_name),
                        estimate_fun = ~style_sigfig(.x,3)) |>
          modify_table_body(
            ~ .x |>
              mutate(estimate_2 = paste0(style_sigfig(estimate,3)," (",ci,")")) |>
              relocate(c(n_obs,estimate_2),.before = estimate) |>
              select(-c(p.value,estimate,ci,conf.high,conf.low))) |>
          modify_header(estimate_2 = "**`Beta (95% CI)`**",N_obs="**N**",label=label)),
    tab_spanner  = model_years)
}

```


# Sensitivity analyses and supplementary results

## Adjust main regression models for ethnicity

```{r ethnicity_models}
# fit fully adjusted model further adjusted for ethnicity
ethnicity_models <- setNames(map(outcome_variables,~geelm(outcome = .x,cov_set="ethnicity"),.progress = T),
                           model_years)

# get table
merge_tables_by_year(ethnicity_models,model_name = "ethnicity")
```

## Community acquired infections: redefine severe infections as:

 1. At least one gp recorded infection diagnostic code on the date of hospital admission or within 7 days prior
 2. Any dispensed antimicrobial on the date of hospital admission or within 7 days prior
 3. Both 1 and 2

```{r community_acquired_infections_exposure}

# approach : exclude setid(matched sets) where the exposed individual does not have the community aquired infection
inf_frailty_cprd |>
  collect() |>
  group_by(setid) |>
  filter(any(gp_infection_7dy_lookback == 1 & exposed ==1)) |> 
  ungroup() |>
  View()


# Fit the fully adjusted regression model on the above data for each comm-acquired inf definition

#  1. At least one outpatient infection diagnostic code on the date of hospital admission or within 7 days prior

s1a_comm_inf_models <- setNames(map(outcome_variables,~geelm(outcome = .x,cov_set="underlying_health",
                                                    data = inf_frailty_cprd |>
                                                      group_by(setid) |>
                                                      filter(any(gp_infection_7dy_lookback == 1 & exposed ==1)) |>
                                                      ungroup())),
                             model_years)

# 2. Any dispensed antimicrobial on the date of hospital admission or within 7 days prior

s1b_comm_inf_models <- setNames(map(outcome_variables,~geelm(outcome = .x,cov_set="underlying_health",
                                                    data = inf_frailty_cprd |>
                                                      group_by(setid) |>
                                                      filter(any(antibio_7dy_lookbck == 1 & exposed ==1)) |>
                                                      ungroup())),
                             model_years)

#   3. The exposed individual has both 1 and 2

#
s1c_comm_inf_models <-  setNames(map(outcome_variables,~geelm(outcome = .x,cov_set="underlying_health",
                                                    data = inf_frailty_cprd |>
                                                      group_by(setid) |>
                                                      filter(any(antibio_7dy_lookbck == 1 & antibio_7dy_lookbck == 1 & exposed ==1)) |>
                                                      ungroup())),
                             model_years)
```

## Restrict the study to periods with likely more frequent recording of some of the deficits included in the eFI due to changes to the Quality Outcomes Framework (QOF) indicators and other policy changes over time; 

1) 2012 onwards 
2) 2014 onwards 

### 1) 2012 onwards 

```{r sensitivity_2012}

#
sensitivity_2012_models <- setNames(map(outcome_variables,~geelm(outcome = .x,cov_set="underlying_health",
                                                          data = inf_frailty_cprd |> 
                                                            filter(indexdate >= as.Date("2012-01-01"))),.progress = T),
                           model_years)

# get table
merge_tables_by_year(sensitivity_2012_models,model_name = "cohort entry from 2012 onwards")
```

### 2) 2014 onwards 
```{r sensitivity_2014}

#
sensitivity_2014_models <- setNames(map(outcome_variables,~geelm(outcome = .x,cov_set="underlying_health",
                                                          data = inf_frailty_cprd |> filter(indexdate >= as.Date("2014-01-01"))),.progress = T),
                           model_years)

# get table
merge_tables_by_year(sensitivity_2014_models,model_name = "cohort entry from 2014 onwards")
```

## Secondary analyses Effect modification (EM) models

Fit models to explore effect modification by:

- Age group
- Deprivation
- Diabetes
- Dementia
- Care home residency
- Severe infection (exposure) type (bacterial, viral, fungal, parasitic)
- Severe infection (exposure) site (sepsis,uti,ssti,mening_enceph,lrti,gastro)

Approach:

- Fit stratified regression models; for each level of EM variable
- Conduct interaction test to determine difference in estimates by level of EM variable. Linear hypothesis F test.

For each EM var:

  - Create a stratified data set for each level
  - Fit the regression models (Years 1 to 5) for each stratified data set
  - Bundle the stratified models into a named list

```{r nest_data_by_interaction_var,eval=TRUE}

# function to stratify data and bundle into a list: could create a list column with nest_by() but too slow

stratify_df <- function(interaction_var,
                        outcome_variables = c("score_1y", "score_2y" ,"score_3y", "score_4y", "score_5y"),
                        model_years =  c("Year 1", "Year 2", "Year 3", "Year 4", "Year 5"),
                        data=inf_frailty_cprd){

  # create sub-cohorts of the em variable
  # for each sub-cohort of em variable, there will be 5 models (one for each ff-up year)
  # to use map2(), will need a .x= data and .y=outcome combination 
  # the number of combinations equals num_levels_interaction_var * num_outcome_vars which is 5

  levels_interaction_var <- data |> pull({{interaction_var}}) |>  factor() |> levels()
  list_data <- map(levels_interaction_var, ~filter(data,{{interaction_var}} == .x))
  list_data <- rep(list_data,length(outcome_variables)) # replicate the data 5 times
  # name the models by Year and level, but sorted in Years so Y1 (levels), Y2 (levels)
  em_model_names <- unlist(map(model_years,~paste0(.x,":",as_label(enquo(interaction_var))," ",levels_interaction_var))) 
  list_data <- setNames(list_data,em_model_names)
}
```

### Fit stratified EM models: fully adjusted models
```{r em_age}
# age
em_models_age <- map2(.x=stratify_df(age_bl_cat4),.y=sort(rep(outcome_variables,4)),
     .f = ~geelm(.y,cov_set = "underlying_health",remove_var = "age_bl_cat4",data = .x),
     .progress = T) 

```

```{r em_sex}
# age
em_models_sex <- map2(.x=stratify_df(gender),.y=sort(rep(outcome_variables,2)),
     .f = ~geelm(.y,cov_set = "underlying_health",remove_var = "gender",data = .x),
     .progress = T) 

```

```{r em_deprivation}
# deprivation
em_models_deprivation <- map2(.x=stratify_df(e2019_imd),.y=sort(rep(outcome_variables,5)),
     .f = ~geelm(.y,cov_set = "underlying_health",remove_var = "e2019_imd",data = .x),
     .progress = T) 

```

```{r em_care_home}

# care home residency
em_models_care_home <- map2(.x=stratify_df(care_home_residency),.y=sort(rep(outcome_variables,2)), 
     .f = ~geelm(.y,cov_set = "underlying_health",data = .x),
     .progress = T)

```

```{r em_dementia}

# dementia
em_models_dem <- map2(.x=stratify_df(dementia),.y=sort(rep(outcome_variables,2)), 
     .f = ~geelm(.y,cov_set = "underlying_health",data = .x),.progress = T)
```

```{r em_diabetes}

# diabetes   
em_models_diab <- map2(.x=stratify_df(diabetes),.y=sort(rep(outcome_variables,2)), 
     .f = ~geelm(.y,cov_set = "underlying_health",data = .x),
     .progress = T)
```

```{r em_inf_type}
# type of severe infection of the exposed individual
em_models_inf_type <-  map2(.x=stratify_df(inf_type_of_match_grp),.y=sort(rep(outcome_variables,4)),
    .f = ~geelm(.y,cov_set = "underlying_health",data = .x),
    .progress = T)
```


```{r em_inf_site}
# site of severe infection of the exposed individual
em_models_inf_site <-  map2(.x=stratify_df(inf_site_of_match_grp),.y=sort(rep(outcome_variables,6)),
    .f = ~geelm(.y,cov_set = "underlying_health",data = .x),
    .progress = T)
```

```{r merge_stack_em_model_tables_by_year_function,eval=TRUE}

# merge yearly em models (for each em var): each year-model summary as a column
# each model(level of em var) as a row

merge_stack_em_tables_by_year <- function(models,levels_interaction_var,label=label){
  stopifnot(is.list(models)) # assert models are supplied as a list

# group models by level of em var
  model_names <- paste0(levels_interaction_var)
  models <- map(levels_interaction_var, ~models[str_detect(names(models),paste0("Year.*[:].*",.x))]) 
  models <- setNames(models,model_names)

  #
  tbls <- imap(models,~merge_tables_by_year(.x,model_name = .y,label = label))
  #
  tbl_stack(tbls)
  
}

```

### Interaction p.values

fit_interaction_model |> conduct_interaction_test |> extract_interaction_p.value

```{r interaction_p_linearhypothesis,eval=TRUE}
p_value_interaction <- function(interaction_model) {
  
  interaction_terms <- matchCoefs(interaction_model,":") 
  n_coefs <- length(interaction_terms) 
  
  # if n_coefs == 1, the null hypothesis is coef == 0 # this pvalue is basically same as pvalue from model (with wald test)
  if(n_coefs ==1) {
    linearHypothesis(interaction_model,interaction_terms,test="F")$`Pr(>F)`[2] |>
      formatC(format = "f",digits=4) 
    
  # if n_coefs > 1, the null hypothesis is that the coefs are equal to each other so,
  # so for 3 coefs, null = coef1 = coef2, coef1=coef3, and coef2=coef3 is implied
    
    } else
      
    if(n_coefs > 1) { 
      linearHypothesis(interaction_model,
                       paste0(interaction_terms[1],
                              "=",
                              interaction_terms[-1]),
                       test="F")$`Pr(>F)`[2] |>
        formatC(format = "f",digits=4) 
    
    }
}

```

```{r interaction_p.values}

# age

em_pval_age <- setNames(
  map(outcome_variables,~geelm(outcome = .x,cov_set="underlying_health",interaction_var = "age_bl_cat4")),
                           model_years) |> 
  map_chr(p_value_interaction) 

# sex

em_pval_sex <- setNames(
  map(outcome_variables,~geelm(outcome = .x,cov_set="underlying_health",interaction_var = "gender")),
                           model_years) |> 
  map_chr(p_value_interaction) 

# deprivation
em_pval_deprivation <- setNames(
  map(outcome_variables,~geelm(outcome = .x,cov_set="underlying_health",interaction_var = "e2019_imd")),
                           model_years) |> 
  map_chr(p_value_interaction) 

# Care home residency
em_pval_care_home <- setNames(
  map(outcome_variables,
      ~geelm(outcome = .x,cov_set="underlying_health",interaction_var = "care_home_residency")),
                           model_years) |> map_chr(p_value_interaction)

# dementia
em_pval_dem <- setNames(
  map(outcome_variables,
      ~geelm(outcome = .x,cov_set="underlying_health",interaction_var = "dementia")),
                           model_years) |> map_chr(p_value_interaction)

# diabetes complicated 
em_pval_diab <- setNames(
  map(outcome_variables,
      ~geelm(outcome = .x,cov_set="underlying_health",interaction_var = "diabetes")),
                           model_years) |> map_chr(p_value_interaction)

# type of severe infection 
em_pval_inf_type <- setNames(
  map(outcome_variables,~geelm(outcome = .x,cov_set="underlying_health",
                               interaction_var = "inf_type_of_match_grp",
                               data = inf_frailty_cprd|> filter(!is.na(inf_type_of_match_grp)))),
                             model_years) |> map_chr(p_value_interaction)

# site of severe infection 
em_pval_inf_site <- setNames(
  map(outcome_variables,~geelm(outcome = .x,cov_set="underlying_health",
                               interaction_var = "inf_site_of_match_grp",
                               data = inf_frailty_cprd|> filter(!is.na(inf_site_of_match_grp)))),
                             model_years) |> map_chr(p_value_interaction)
# combine p values
em_pvalues <- bind_rows(em_pval_age,em_pval_sex,em_pval_deprivation,
                        em_pval_care_home,em_pval_dem,em_pval_diab,
                        em_pval_inf_type,em_pval_inf_site) |>
  tibble() |>
  mutate(variable = c("Age","Sex","Deprivation","Care home residency","Dementia","Diabetes","Type of severe infection","Site of severe infection"),.before=1)

```

## Prevalence of eFI deficits over time
```{r efi_deficits_overtime}

#
efi_deficit_names <- sort(unique(str_replace(names(collect(eFI_deficits_data)),"_(bl|yr[1-5])","")))

efi_deficit_names_sorted <- Filter(function(x) !x %in% c("setid","enddate","patid","indexdate"),efi_deficit_names)

#
efi_deficits_bl <- collect(eFI_deficits_data) |> 
  filter(if_any(matches("_yr[1-5]$"),~!is.na(.x))) |>
  filter(if_all(matches("_yr1$"),~!is.na(.x))) |> # keep as yr1: those censored at entry do not have yr1
  select(matches("_bl$"),patid,setid) |> 
  rename_with(~str_replace(.x,"_..$","")) |>
  select(any_of(efi_deficit_names_sorted),patid,setid) |>
  filter(if_all(matches("_bl$"),~!is.na(.x))) |>
  left_join(select(collect(cohort_matched_2),patid,setid,exposed),
            join_by(patid,setid)) |>
  tbl_summary(by=exposed,include = -c(patid,setid),missing = "no")

#
efi_deficits_yr1 <- collect(eFI_deficits_data) |>
  filter(if_any(matches("_yr[1-5]$"),~!is.na(.x))) |>
  filter(if_all(matches("_yr1$"),~!is.na(.x))) |>
  select(matches("_yr1$"),patid,setid) |>
  rename_with(~str_replace(.x,"_...$","")) |>
  select(any_of(efi_deficit_names_sorted),patid,setid) |>
  left_join(select(collect(cohort_matched_2),patid,setid,exposed),
            join_by(patid,setid)) |>
  tbl_summary(by=exposed,include = -c(patid,setid),missing = "no")

#
efi_deficits_yr2 <- collect(eFI_deficits_data) |>
  filter(if_any(matches("_yr[1-5]$"),~!is.na(.x))) |>
  filter(if_all(matches("_yr2$"),~!is.na(.x))) |>
  select(matches("_yr2$"),patid,setid) |> 
  rename_with(~str_replace(.x,"_...$","")) |>
  select(any_of(efi_deficit_names_sorted),patid,setid) |>
  left_join(select(collect(cohort_matched_2),patid,setid,exposed),
            join_by(patid,setid)) |>
  tbl_summary(by=exposed,include = -c(patid,setid),missing = "no")

#
efi_deficits_yr3 <- collect(eFI_deficits_data) |>
  filter(if_any(matches("_yr[1-5]$"),~!is.na(.x))) |>
  filter(if_all(matches("_yr3$"),~!is.na(.x))) |>
  select(matches("_yr3$"),patid,setid) |>
  rename_with(~str_replace(.x,"_...$","")) |>
  select(any_of(efi_deficit_names_sorted),patid,setid) |>
  left_join(select(collect(cohort_matched_2),patid,setid,exposed),
            join_by(patid,setid)) |>
  tbl_summary(by=exposed,include = -c(patid,setid),missing = "no")

#
efi_deficits_yr4 <- collect(eFI_deficits_data) |>
  filter(if_any(matches("_yr[1-5]$"),~!is.na(.x))) |>
  filter(if_all(matches("_yr4$"),~!is.na(.x))) |>
  select(matches("_yr4$"),patid,setid) |>
  rename_with(~str_replace(.x,"_...$","")) |>
  select(any_of(efi_deficit_names_sorted),patid,setid) |>
  left_join(select(collect(cohort_matched_2),patid,setid,exposed),
            join_by(patid,setid)) |>
  tbl_summary(by=exposed,include = -c(patid,setid),missing = "no")

#
efi_deficits_yr5 <- collect(eFI_deficits_data) |>
  filter(if_any(matches("_yr[1-5]$"),~!is.na(.x))) |>
  filter(if_all(matches("_yr5$"),~!is.na(.x))) |>
  select(matches("_yr5$"),patid,setid) |>
  rename_with(~str_replace(.x,"_...$","")) |>
  select(any_of(efi_deficit_names_sorted),patid,setid) |>
  left_join(select(collect(cohort_matched_2),patid,setid,exposed),
            join_by(patid,setid)) |>
  tbl_summary(by=exposed,include = -c(patid,setid),missing = "no")

# 
tbl_merge(tbls=list(efi_deficits_bl,
                    efi_deficits_yr1,
                    efi_deficits_yr2,
                    efi_deficits_yr3,
                    efi_deficits_yr4,
                    efi_deficits_yr5))
```

