---
title: "A multistate model of frailty progression after severe infections in adults â‰Ą65 years in England: a matched-cohort study"
author: "Kwabena Asare"
date: "`r lubridate::today()`"
format: html
execute: 
  include: true
  echo: false
  eval: false
editor_options: 
  chunk_output_type: console
---

## Main Analyses

**Individuals making at least one follow-up**: this is the main analysis sample: dataset already filtered accordingly

### Table 1: Descriptive results at indexdate

**Table 1.** Descriptive summaries of **baseline** characteristics by severe infection (exposure) status.

```{r table_one_main}

# baseline characteristics by exposure status

table_one_main <- collect(inf_frailty_cprd) |>
  transmute(exposed = if_else(exposed == 0, "No","Yes"),
            age_bl,age_bl_cat4,gender,ethnicity,
            e2019_imd,harmful_alcohol_use,
            smoking,care_home_residency,
            gp_infection_5yr_lookback,gp_infection_7dy_lookback,
            antibio_7dy_lookbck,dementia,diabetes,
            atrial_fibrillation_bl,
            chronic_kidney_disease_bl,hypertension_bl,
            urinary_system_disease_bl,
            osteoporosis_bl,parkinsonism_tremors_bl,
            ischaemic_heart_disease_bl,heart_failure_bl,
            cerebrovascular_disease_bl,rheumatoid_diseases_bl,
            eFI_bl,eFI_bl_cat,one_consult_1y_lookbck) |> collect() |>
  rename_with(.fn = ~str_replace(.x,"_bl","")) |>
  tbl_summary(by="exposed",
              missing_text = "Missing",
              type=list(c(age,eFI) ~ "continuous2"),
              statistic = list(all_continuous2() ~ c("{mean} ({sd})","{median} ({p25}, {p75})")),
              digits = list(all_continuous() ~ 0,c(eFI) ~ 2)) |>
  bold_labels() |>
  modify_footnote(everything() ~ NA) |>
  modify_caption(glue("**Table 1.** Characteristics of participants at index date (England)")) |>
  modify_spanning_header(all_stat_cols() ~ "**Severe infection after age 65 years**")

table_one_main
```

## Multi-state survival analysis: Cox regression
## Transition specific Cox regression models

For each outcome build models with the following sequence of confounder adjustment sets:

 - **Minimally adjusted** models: adjust for matching variables (age, gender), ethnicity: 
   **No need to adjust for baseline eFI category since this is same for each transition risk set**
   **However can adjust for the eFI score though**
 - **Deprivation adjusted** models: further adjust minimal model for:
    - individual and practice level IMD (quntiles)
 - **Lifestyle adjusted** models: further adjust deprivation model for:
    - smoking
    - harmful alcohol use
 - **Infection history** models: further adjust lifestyle model for:
  - 5 yr look back infections; gp recorded

### Step 1: Set up transition matrix
 - The transition matrix specifies which direct transitions are possible
(those with NA are impossible) and assigns numbers to the transitions for future reference.
 - **Note:** Only forward transitions are possible since eFI 1.0 only ever increases

```{r transMat}
# Done manually

tmat_inf_frail <- matrix(NA,5,5)

dimnames(tmat_inf_frail) <- list(from = c("fit", "mild_frail", "mod_frail","severe_frail","death"),
                       to = c("fit", "mild_frail", "mod_frail","severe_frail","death"))

tmat_inf_frail[1,2:5] <- 1:4
tmat_inf_frail[2,3:5] <- 5:7
tmat_inf_frail[3,4:5] <- 8:9
tmat_inf_frail[4,5] <- 10

# With transMat() from mstate package
transMat(x=c(list(c(2:5),c(3:5),c(4:5),c(5),c())),
         names = c("fit", "mild_frail", "mod_frail","severe_frail","death"))

```

### Step 2: Prepare data for mstate

```{r reshape_mstate}

# covariates: condounder adjustment set categories
matching <- c("exposed","eFI_bl") 
deprivation <- c("e2019_imd")
lifestyle <- c("harmful_alcohol_use","smoking")
infection_history <- c("gp_infection_5yr_lookback")
ethnicity <- c("ethnicity") 
covariates <- c(matching,deprivation,lifestyle,infection_history)
other_covariates <- c("patid","setid","indexdate","gp_infection_7dy_lookback","antibio_7dy_lookbck","diabetes","dementia",
                        "care_home_residency","inf_site_of_match_grp","inf_type_of_match_grp")
  
# mstate data
# THOSE THAT DO NOT START FROM FIT BUT A HIGHER STATE, THEY CANNOT MOVE TO THEIR STARTING STATES BUT ONLY MOVE FROM THEM
# SO AFTER PREPING THE DATA REMOVE ALL THEIR TRANSITIONS BEFORE THEIR STARTING STATES
# ALL SUCH TRANSITIONS WILL HAVE STOP TIME == 0
inf_frailty_cprd_mstate <- msprep(time = c(NA,"eFI_mild_years","eFI_moderate_years","eFI_severe_years","eFI_death_years"),
                status = c(NA,"eFI_mild","eFI_moderate","eFI_severe","eFI_death"), 
                data = left_join(select(collect(eFI_mstate_data),-c(enddate,indexdate)),
                                 select(collect(inf_frailty_cprd),any_of(c(covariates,ethnicity,other_covariates))),
                                 join_by(patid,setid)) |> mutate(setid=as.numeric(factor(setid))), 
                trans = tmat_inf_frail,
                keep = c(covariates,other_covariates,ethnicity)) |>
  # remove transitions into non eFI_fit starting states
   filter(Tstop!=0) 

# distribution of events
 events_all <- events(inf_frailty_cprd_mstate) 

# in the exposed (severe infection)
 events_exposed <- events(filter(inf_frailty_cprd_mstate,exposed==1))
# in the unexposed
 events_un_exposed <- events(filter(inf_frailty_cprd_mstate,exposed==0))
```


### Step 3: Add transition-specific covariates to the data
```{r transition_specific_covariates}

# data with all expanded covariates
inf_frailty_cprd_mstate_expcovs <- expand.covs(inf_frailty_cprd_mstate,
                                               covs = c(covariates,ethnicity),append = T,longnames = F) |>
  to_duckdb(con = my_con_integer64)

#expanded covariate sets: 
expanded_covariates  <-  grep("[.]\\d+$",names(select(collect(inf_frailty_cprd_mstate_expcovs),matches(glue("{covariates}.*")))),value = T)
# matching factors  
expanded_covariates_matching  <- grep("[.]\\d+$",names(select(collect(inf_frailty_cprd_mstate_expcovs),matches(glue("{matching}.*")))),value = T)

expanded_covariates_deprivation  <-   grep("[.]\\d+$",names(select(collect(inf_frailty_cprd_mstate_expcovs),matches(glue("{c(matching,deprivation)}.*")))),value = T)

expanded_covariates_lifestyle  <-  grep("[.]\\d+$",names(select(collect(inf_frailty_cprd_mstate_expcovs),matches(glue("{c(matching,deprivation,lifestyle)}.*")))),value = T)

expanded_covariates_infection_history <-  grep("[.]\\d+$",names(select(collect(inf_frailty_cprd_mstate_expcovs),matches(glue("{c(matching,deprivation,lifestyle,infection_history)}.*")))),value = T)

expanded_covariates_ethnicity <-  grep("[.]\\d+$",names(select(collect(inf_frailty_cprd_mstate_expcovs),matches(c(glue("{c(matching,deprivation,lifestyle,infection_history)}.*"),"ethnicity")))),value = T)

```

### Step 3 Estimation
#### Transition stratified and setid stratified Cox regression to obtain all transition-specific HR and transition probabilities from one model

```{r all_mstate_cox_matching}
# NOTE: to include only covariate sets not missing
# Since these are levels in the data ("Missing")

all_mstate_cox_matching <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_matching,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = collect(inf_frailty_cprd_mstate_expcovs))
```

```{r all_mstate_cox_deprivation}
all_mstate_cox_deprivation <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_deprivation,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),
                                               if_all(all_of(c(matching,deprivation)),~(!is.na(.x) & .x != "Missing"))))
```

```{r all_mstate_cox_lifestyle}
all_mstate_cox_lifestyle <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_lifestyle,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),
                                               if_all(all_of(c(matching,deprivation,lifestyle)),~(!is.na(.x) & .x != "Missing"))))
```

```{r all_mstate_cox_infection_history}
all_mstate_cox_infection_history <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
```


#### Figure 2: Forest plots of fully adjusted transition specific HR (95% CI): Severe infections vs no severe infection

```{r forest_plot}
# data.frame one: extract model elements into a dataframe; HRs and 95% CIs
forestplot_mstate_HRs <- function(model, # a cox model with all transiotioin HRs
                            events_exposed, # events in the exposed; obtained with mstate::events()
                            events_un_exposed,
                            by_sub_group=T,
                            clip=5,legend_x_pos=0.01,legend_y_pos=0.01,legend_dist=0.195,plot_mar=5,
                            plot_save_name,height=15,width=28,...) {

  plot_df <- if(is.data.frame(model)) model else 
    tidy(model,exponentiate = T,conf.int = T)[1:10,c("term","estimate", "conf.low","conf.high")] |>
      # Add n. events/at risk number of events and N
    mutate(events_exp = c(events_exposed$Frequencies[1,2],events_exposed$Frequencies[1,3],
                          events_exposed$Frequencies[1,4],events_exposed$Frequencies[1,5],
                          events_exposed$Frequencies[2,3],events_exposed$Frequencies[2,4],
                          events_exposed$Frequencies[2,5],events_exposed$Frequencies[3,4],
                          events_exposed$Frequencies[3,5],events_exposed$Frequencies[4,5]),
           sample_size_exp = c(rep(events_exposed$Frequencies[1,7],4),
                               rep(events_exposed$Frequencies[2,7],3),
                               rep(events_exposed$Frequencies[3,7],2),
                               events_exposed$Frequencies[4,7])) |>
    mutate(events_un_exp = c(events_un_exposed$Frequencies[1,2],events_un_exposed$Frequencies[1,3],
                             events_un_exposed$Frequencies[1,4],events_un_exposed$Frequencies[1,5],
                             events_un_exposed$Frequencies[2,3],events_un_exposed$Frequencies[2,4],
                             events_un_exposed$Frequencies[2,5],events_un_exposed$Frequencies[3,4],
                             events_un_exposed$Frequencies[3,5],events_un_exposed$Frequencies[4,5]),
           sample_size_un_exp = c(rep(events_un_exposed$Frequencies[1,7],4),
                                  rep(events_un_exposed$Frequencies[2,7],3),
                                  rep(events_un_exposed$Frequencies[3,7],2),
                                  events_un_exposed$Frequencies[4,7])) |>
    mutate(n_N_exp = paste0(events_exp,"/",sample_size_exp),
           n_N_un_exp = paste0(events_un_exp,"/",sample_size_un_exp)) |>
    # Mast events from 1:50
    mutate(n_N_exp = if_else(events_exp %in% 1:50,paste0("<50","/",sample_size_exp),n_N_exp),
           n_N_un_exp = if_else(events_un_exp %in% 1:50,paste0("<50","/",sample_size_un_exp),n_N_un_exp)) |>
    mutate(term_description = c("Fit to Mild Frailty","Fit to Moderate Frailty",
                                "Fit to Severe Frailty","Fit to Death",
                                "Mild to Moderate Frailty","Mild to Severe Frailty",
                                "Mild Frailty to Death","Moderate to Severe Frailty",
                                "Moderate Frailty to Death","Severe Frailty to Death"),.after = 1) |>
    mutate(term_description = factor(term_description,
                                     levels = c("Fit to Mild Frailty","Fit to Moderate Frailty",
                                                "Fit to Severe Frailty","Fit to Death",
                                                "Mild to Moderate Frailty","Mild to Severe Frailty",
                                                "Mild Frailty to Death","Moderate to Severe Frailty",
                                                "Moderate Frailty to Death","Severe Frailty to Death"))) |>
    mutate(across(c(conf.low,conf.high),~if_else(.x < 0.0001 | .x > 100,0,.x))) |>
    mutate(est_label = if_else(is.na(estimate),NA_character_,
                               paste0(formatC(estimate,format = "f",digits = 2),
                                                                    " (",formatC(conf.low,format = "f",digits = 2),", ",
                                                                    formatC(conf.high,format = "f",digits = 2),")")),
           across(c(estimate,conf.low,conf.high),~round(.x,2)) ) |>
    #filter(term_description != "Fit to Severe Frailty") |>
    mutate(across(c(estimate,conf.low,conf.high),~if_else(.x < 0.0001 | .x > 100,NA,.x))) |>
    mutate(est_label_middot=paste0(str_replace_all(est_label,"[.]","\U00B7")),
           estimate_middot = str_replace(estimate,"[.]","\U00B7")) |>
    mutate(across(c(conf.low,conf.high,est_label,estimate),~ if_else(events_exp==0|events_un_exp==0,NA,.x))) |>
    mutate(sub_group=0)
  
  # graphical parameters
  plot_cols <- c("darkblue","darkred","darkgreen","orange","purple","yellow4")
  
  line_cols <-  case_when(as.numeric(plot_df$sub_group)==1~plot_cols[[1]],
                          as.numeric(plot_df$sub_group)==2 ~plot_cols[[2]],
                          as.numeric(plot_df$sub_group)==3 ~plot_cols[[3]],
                          as.numeric(plot_df$sub_group)==4 ~plot_cols[[4]],
                          as.numeric(plot_df$sub_group)==5 ~plot_cols[[5]],
                          as.numeric(plot_df$sub_group)==6 ~plot_cols[[6]],.default = "black")

    
    #  forest plot
             fp <- forestplot(plot_df,
             labeltext=c(term_description,n_N_exp,n_N_un_exp,est_label_middot),
             mean = estimate,
             lower = conf.low,
             upper = conf.high,
             clip = c(-Inf,clip),
             xticks =seq(0,clip,1),
             xlab="HR (95% CI)",
             boxsize=.20,
             zero=1,
             mar = unit(rep(plot_mar, times = 4), "mm")) |>  # bottom, left, top, right) 
    fp_add_header(n_N_exp = "Severe infection\n(n. events/at risk)",
                  n_N_un_exp = "No severe infection\n(n. events/at risk)",
                  est_label_middot= "Adjusted HR (95% CI)") |>
    fp_set_style(line = if(isTRUE(by_sub_group)) c("black",line_cols) else "black",
                 box = if(isTRUE(by_sub_group)) c("black",line_cols) else "black",
                 zero =gpar(col = "black", lty = "solid")) |>
    fp_decorate_graph(graph.pos = 4) |>
    fp_add_lines(h_2 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),...)

 
  # save forest plots
  saveRDS(fp,glue("{getwd()}/results/{plot_save_name}.rds"))
  
  png(glue("{getwd()}/results/{plot_save_name}.png"),height = height,width = width,units = "cm",res = 300) # 25cm = 250mm, min is 107mm, 300 dpi resolution
  if(isFALSE(by_sub_group)) print(fp) else { 
    print(fp)
      legend_labels <- levels(plot_df$sub_group)
      legend_cols   <- plot_cols[1:length(levels(plot_df$sub_group))]
      
      for (i in seq_along(legend_labels)) {
        
        grid.points(
          y = unit(legend_y_pos, "npc"),
          x = unit(legend_x_pos+i*legend_dist, "npc"),
          pch = 15,
          gp = gpar(col = legend_cols[i])
        )
        
        grid.text(
          legend_labels[i],
          y = unit(legend_y_pos, "npc"),
          x = unit(legend_x_pos+0.007+i*legend_dist, "npc"),
          just = "left",
          gp = gpar(fontsize = 12,fontface="bold")
        )
      }
    }
  
  dev.off()
  
  tiff(glue("{getwd()}/results/{plot_save_name}.tiff"),height = height,width = width,units = "cm",res = 300) # 25cm = 250mm, min is 107mm, 300 dpi resolution
  if(isFALSE(by_sub_group)) print(fp) else { 
    print(fp)
      legend_labels <- levels(plot_df$sub_group)
      legend_cols   <- plot_cols[1:length(levels(plot_df$sub_group))]
      
      for (i in seq_along(legend_labels)) {
        
        grid.points(
          y = unit(legend_y_pos, "npc"),
          x = unit(legend_x_pos+0.007+i*legend_dist, "npc"),
          pch = 15,
          gp = gpar(col = legend_cols[i])
        )
        
        grid.text(
          legend_labels[i],
          y = unit(legend_y_pos, "npc"),
          x = unit(legend_x_pos+0.007+i*legend_dist, "npc"),
          just = "left",
          gp = gpar(fontsize = 12,fontface="bold")
        )
      }
    }
  
  dev.off()
  
}

```

```{r forest_plot_full_model}
  # forest plot: main
forestplot_mstate_HRs(model = all_mstate_cox_infection_history.,
                                            events_exposed = events(filter(inf_frailty_cprd_mstate,
                                                                           exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                                            events_un_exposed = events(filter(inf_frailty_cprd_mstate,
                                                                           exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                      by_sub_group = F,
                                            plot_save_name = "Figure_3_mstate",
                      h_6 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_9 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_11 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2))
```

### Figure_S2: Forest plots sequential adjustments
## Create a function to extract and combine model data from sub-group analyses

```{r extract_model_data}

extract_model_data <- function(model,events_exposed,events_un_exposed,sub_group_level=NULL,sub_group_desc=NULL) { 
  
  model <- if(is.data.frame(model)) model else tidy(model,exponentiate = T,conf.int = T)[1:10,c("term","estimate", "conf.low","conf.high")] 
  
  
  plot_df <- model|> 
      # Add n. events/at risk number of events and N
    mutate(events_exp = c(events_exposed$Frequencies[1,2],events_exposed$Frequencies[1,3],
                        events_exposed$Frequencies[1,4],events_exposed$Frequencies[1,5],
                        events_exposed$Frequencies[2,3],events_exposed$Frequencies[2,4],
                        events_exposed$Frequencies[2,5],events_exposed$Frequencies[3,4],
                        events_exposed$Frequencies[3,5],events_exposed$Frequencies[4,5]),
         sample_size_exp = c(rep(events_exposed$Frequencies[1,7],4),
                             rep(events_exposed$Frequencies[2,7],3),
                             rep(events_exposed$Frequencies[3,7],2),
                             events_exposed$Frequencies[4,7])) |>
  mutate(events_un_exp = c(events_un_exposed$Frequencies[1,2],events_un_exposed$Frequencies[1,3],
                           events_un_exposed$Frequencies[1,4],events_un_exposed$Frequencies[1,5],
                           events_un_exposed$Frequencies[2,3],events_un_exposed$Frequencies[2,4],
                           events_un_exposed$Frequencies[2,5],events_un_exposed$Frequencies[3,4],
                           events_un_exposed$Frequencies[3,5],events_un_exposed$Frequencies[4,5]),
         sample_size_un_exp = c(rep(events_un_exposed$Frequencies[1,7],4),
                                rep(events_un_exposed$Frequencies[2,7],3),
                                rep(events_un_exposed$Frequencies[3,7],2),
                                events_un_exposed$Frequencies[4,7])) |>
  mutate(n_N_exp = paste0(events_exp,"/",sample_size_exp),
         n_N_un_exp = paste0(events_un_exp,"/",sample_size_un_exp)) |>
  # Mast events from 1:50
  mutate(n_N_exp = if_else(events_exp %in% 1:50,paste0("<50","/",sample_size_exp),n_N_exp),
         n_N_un_exp = if_else(events_un_exp %in% 1:50,paste0("<50","/",sample_size_un_exp),n_N_un_exp)) |>
  mutate(term_description = c("Fit to Mild Frailty","Fit to Moderate Frailty",
                              "Fit to Severe Frailty","Fit to Death",
                              "Mild to Moderate Frailty","Mild to Severe Frailty",
                              "Mild Frailty to Death","Moderate to Severe Frailty",
                              "Moderate Frailty to Death","Severe Frailty to Death"),.after = 1) |>
  mutate(term_description = factor(term_description,
                                   levels = c("Fit to Mild Frailty","Fit to Moderate Frailty",
                                              "Fit to Severe Frailty","Fit to Death",
                                              "Mild to Moderate Frailty","Mild to Severe Frailty",
                                              "Mild Frailty to Death","Moderate to Severe Frailty",
                                              "Moderate Frailty to Death","Severe Frailty to Death"))) |>
  mutate(across(c(estimate,conf.low,conf.high),~if_else(.x < 0.0001 | .x > 100,NA,.x))) |>
  mutate(est_label = if_else(is.na(estimate),NA_character_,
                             paste0(formatC(estimate,format = "f",digits = 2),
                                    " (",formatC(conf.low,format = "f",digits = 2),", ",
                                    formatC(conf.high,format = "f",digits = 2),")")),
         across(c(estimate,conf.low,conf.high),~round(.x,2)) ) |>
  #filter(term_description != "Fit to Severe Frailty") |>
  mutate(est_label_middot=paste0(str_replace_all(est_label,"[.]","\U00B7")),
         estimate_middot = str_replace(estimate,"[.]","\U00B7")) |>
  mutate(across(c(conf.low,conf.high,est_label,estimate),~ if_else(events_exp==0|events_un_exp==0,NA,.x))) |>
    mutate(sub_group = factor(sub_group_level,labels=sub_group_desc))
  }


```


```{r forest_plot_sequential_adjustment}

plot_df_sequential_adjustment <- rbind( # matching
  extract_model_data(model = all_mstate_cox_matching,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,exposed==1)),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,exposed==0)),sub_group_level = 1,sub_group_desc="Baseline frailty"),
  # deprivation
  extract_model_data(model = all_mstate_cox_deprivation,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,exposed==1,if_all(all_of(covariates),~(!is.na(.x) & .x != "Missing")))),
                      events_un_exposed = events(filter(inf_frailty_cprd_mstate,exposed==0,if_all(all_of(covariates),~(!is.na(.x) & .x != "Missing")))),sub_group_level = 2,sub_group_desc="Deprivation"),
  # lifestyle
  extract_model_data(model = all_mstate_cox_lifestyle,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,exposed==1,if_all(all_of(c(covariates,deprivation,lifestyle)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,exposed==0,if_all(all_of(c(covariates,deprivation,lifestyle)),~(!is.na(.x) & .x != "Missing")))),sub_group_level = 3,sub_group_desc="Lifestyle"),
  # infection_history
  extract_model_data(model = all_mstate_cox_infection_history.,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,exposed==1,if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,exposed==0,if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),sub_group_level = 4,sub_group_desc="Infection histroy")) |>
  arrange(term_description,sub_group) |>
  mutate(term_description = if_else(duplicated(term_description),NA_character_,term_description)) 

# Plot
forestplot_mstate_HRs(model = plot_df_sequential_adjustment,
                      height = 45,width = 32,
                      plot_save_name = "Figure_S2_mstate",
                      h_6 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_10 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_14 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_18 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_22 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_26 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_30 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_34 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_38 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2))
```


### Step 4: Prediction

This is basically prediction of the cumulative incidence.
Just to highlight that for CAUSE SPECIFIC HAZARDS which we have estimated for the multi-state modeling,
effects on the cumulative incidence function may be different from what the regression coefficients (HR) suggest.
AKA, no one-to-one relationship between the HR and the CI function.

Thus these estimates of the transition probabilities of events are based on the results obtained from the Cox model on the transition hazards
between the states.

What is being estimated here is the (Prob of an event | event history and a set of covariates) at a particular time.
Estimating this over time gives the curve of the probabilities. To specifically estimate these among exposed and unexposed adults:

 1. Probabilities of entering each state and 
 2. Mean state occupation times
 3. **To do up to 5 years based on mean follow-up time**
 4. Only for the fully adjusted models

Given a multi-state model without recurrent events, these probabilities can be expressed in terms
of the hazards for the transitions and can be estimated by appropriately combining the estimated
baseline hazards and regression coefficients.

#### Prediction data
```{r prediction_data}

# New data for one exposed and unexposed person
# newdata data frame needs to have exactly one line for each transition per person
# must be of msdata class with a trans attribute

# un_exp
new_data_un_exp <- data.frame(matrix(rep(0,60),ncol = 6)) |>
  rename_with(~covariates) |>
  mutate(trans=1:10,strata=1:10) |>
  mutate(smoking = factor(smoking,levels = 0:3,
                              labels = unique(collect(inf_frailty_cprd)$smoking)))

attr(new_data_un_exp,"trans") <- tmat_inf_frail
class(new_data_un_exp) <- c("msdata","data.frame")
new_data_un_exp <- expand.covs(new_data_un_exp,covariates,longnames=F)

# exp
new_data_exp <- data.frame(matrix(rep(0,60),ncol = 6)) |>
  rename_with(~covariates) |>
  mutate(trans=1:10,strata=1:10,exposed=1)|>
  mutate(smoking = factor(smoking,levels = 0:3,
                              labels = unique(collect(inf_frailty_cprd)$smoking)))

attr(new_data_exp,"trans") <- tmat_inf_frail
class(new_data_exp) <- c("msdata","data.frame")
new_data_exp <- expand.covs(new_data_exp,covariates,longnames=F)
```

#### Predicted Hazards
```{r predict_hazards}

# Obtain patient specific transition hazards: using the baseline hazards and coefficients from the cox model: to use the msfit() function

# The result is an object of class msfit, which is a list with three items, Haz, varHaz, and trans

# NOTE: for each patient all predictions are at the event times


baseline_hz_predict_un_exp  <- msfit(all_mstate_cox_infection_history., newdata = new_data_un_exp, trans = tmat_inf_frail)

summary(baseline_hz_predict_un_exp)
baseline_hz_predict_un_exp[[1]]

#
baseline_hz_predict_exp  <- msfit(all_mstate_cox_infection_history., newdata = new_data_exp, trans = tmat_inf_frail)

#
summary(baseline_hz_predict_exp)
baseline_hz_predict_exp[[1]]
```

#### Predicted probability of transition
```{r predict_prob_trans}

#' We are interested in estimating so called transition probabilities
#' For instance, P13(0, t) indicates the probability of having been moderately frail (state 3)
#' by time t, say yr5 given that the individual was fit in state (state 1) at time s = 0

# use the resulting patient-specific transition hazards (and variances and covariances) as input
# for probtrans() to obtain (patient-specific) transition probabilities
# Note: one exposed and one unexposed person

#' The result of probtrans is a probtrans object, which is a list, where item [[i]] contains
#' predictions from state i. Each item of the list is a data frame with time containing all event
#' time points, and pstate1, pstate2, etc the probabilities of being in state 1, 2, etc, and finally
#' se1, se2 etc the standard errors of these estimated probabilities.

# NOTE: To stress that the predictions are from a particular state and the probs represent the prob of 
# being in a particular state starting from a particular previous state at that event time
# for a prediction at a particular time, you can use the time var to filter
# Also NOTE that the for each element in the list representing the transition,
# the first col gives the prob of remaining in the state without transitioning to a higher state

# So the plot actually plots the n = trans col; the col corresponding to the transition

# starting from Fit
probtrans_un_exp <- probtrans(baseline_hz_predict_un_exp,predt = 0)

summary(probtrans_un_exp,from =1)
head(probtrans_un_exp[[1]])
slice(filter(probtrans_un_exp[[1]],time>=5),1)

#
probtrans_exp <- probtrans(baseline_hz_predict_exp,predt = 0)
summary(probtrans_exp,from =1)
head(probtrans_exp[[1]])
slice(filter(probtrans_exp[[1]],time>=5),1)
```

```{r state_occupancy_prob_yr5}

# State occupancy probabilities starting from each state at a time
# Death is the absorbing state so can't start from death

state_occu_prob_5 <- rbind(
  slice(filter(probtrans_exp[[1]],time>=5),1),
  slice(filter(probtrans_exp[[2]],time>=5),1),
  slice(filter(probtrans_exp[[3]],time>=5),1),
  slice(filter(probtrans_exp[[4]],time>=5),1),
  slice(filter(probtrans_un_exp[[1]],time>=5),1),
  slice(filter(probtrans_un_exp[[2]],time>=5),1),
  slice(filter(probtrans_un_exp[[3]],time>=5),1),
  slice(filter(probtrans_un_exp[[4]],time>=5),1)) |>
  rename(Fit = pstate1,
         Mild = pstate2,
         Moderate = pstate3,
         Severe = pstate4,
         Death = pstate5,
         Fit_se = se1,
         Mild_se = se2,
         Moderate_se = se3,
         Severe_se = se4,
         Death_se = se5) |>
  mutate(across(Fit:Death_se,~round(.x,3)*100)) |>
  mutate(exposure = c(1,1,1,1,0,0,0,0),
         from = rep(c("Fit","Mild","Moderate","Severe"),2),.after=time) |>
   # calculate 95% CI
  mutate(across(matches("(Fit|Mild|Moderate|Severe|Death)$"),
                ~paste0(.x," (",.x-round(1.96*get(glue("{cur_column()}_se")),1),", ",
                                                      .x+round(1.96*get(glue("{cur_column()}_se")),1),")"),
                .names = "{.col}_ci")) |>
  select(time,exposure,from,matches("ci")) |>
  mutate(across(matches("ci"),~if_else(.x=="0 (0-0)",NA_character_,.x)))

#
states <- c("Fit","Mild","Moderate","Severe","Death")

state_occu_prob_5 |>
  select(-c(time)) |>
  relocate(from) |>
  gt(rowname_col = "from",caption = "Probability of transition at 5 years")
```

#### Figure 3: Stacked predicted state occupancy probabilities over time and mean state occupancy times
```{r stacked_plots_from_fit}

# Plot transition probabilities from state 1; the from arg in the plot.msfit(); default from == 1
# So this plot uses one transition to plot from 
# the estimates are still prob of remaining in a state; starting from state 1 (fit)
# can do other graphs starting from each state

# Step 3: Stacked predicted probability of transitions plot
# The default type of the plot method for probtrans objects is a "stacked" plot, for which the
# difference between two adjacent lines represents the probability of being in a state

# From fit
plot_exp_from_fit <- plot(probtrans_exp,
     ord=c(1:5),linewidth=2,
     xlab = "Years since entering fit frailty state at index date",
     ylab = "Probability",
     cex=0.75,
     xlim = c(0,6),
     main="A. From fit status (severe infection)",
     legend = c("Fit","Mild Frailty","Moderate Frailty","Severe Frailty","Death"),
     legend.pos = "top",
     cols = c("#4DAF4A","#377EB8","#984EA3","#8B8B00","#E41A1C"), # inspired by brewer.pal(5,"Set1"),
     use.ggplot = T) 


plot_un_exp_from_fit <- plot(probtrans_un_exp,
     ord=c(1:5),linewidth=2,
     xlab = "Years since entering fit frailty state at index date",
     ylab = "Probability",
     cex=0.75,
     xlim = c(0,6),
     main="B. From fit status (no severe infection)",
     legend = c("Fit","Mild Frailty","Moderate Frailty","Severe Frailty","Death"),
     legend.pos = "top",
     cols = c("#4DAF4A","#377EB8","#984EA3","#8B8B00","#E41A1C"), # inspired by brewer.pal(5,"Set1"),
     use.ggplot = T)

# combine plots
stacked_plot_from_fit <- ggarrange(plot_exp_from_fit + ggtitle("A. From fit status (no severe infection)"),
                          plot_un_exp_from_fit + ggtitle("B. From fit status (no severe infection)"),
                          common.legend = T,
                          nrow=1,hjust = -0)
```

```{r stacked_plots_from_mild}

# From mild
plot_exp_from_mild <- plot(probtrans_exp,from=2,
     ord=c(2:5),linewidth=2,
     xlab = "Years since entering mild frailty state at/after index date",
     ylab = "Probability",
     cex=0.75,
     xlim = c(0,6),
     main="C. From mild frailty (severe infection)",
     legend = c("Fit","Mild Frailty","Moderate Frailty","Severe Frailty","Death"),
     legend.pos = "top",
     cols = c("#377EB8","#984EA3","#8B8B00","#E41A1C"), # inspired by brewer.pal(5,"Set1"),
     use.ggplot = T)


plot_un_exp_from_mild <- plot(probtrans_un_exp,from=2,
     ord=c(2:5),linewidth=2,
     xlab = "Years since entering mild frailty state at/after index date",
     ylab = "Probability",
     cex=0.75,
     xlim = c(0,6),
     main="D. From mild frailty (no severe infection)",
     legend = c("Fit","Mild Frailty","Moderate Frailty","Severe Frailty","Death"),
     legend.pos = "top",
     cols = c("#377EB8","#984EA3","#8B8B00","#E41A1C"), # inspired by brewer.pal(5,"Set1"),
     use.ggplot = T)

# combine plots
stacked_plot_from_mild <- ggarrange(plot_exp_from_mild + ggtitle("C. From mild frailty (no severe infection)"),
                          plot_un_exp_from_mild + ggtitle("D. From mild frailty (no severe infection)"),
                          common.legend = T,
                          nrow=1,hjust = -0)
  
```

```{r stacked_plots_from_moderate}

# From mild
plot_exp_from_moderate <- plot(probtrans_exp,from=3,
     ord=c(3:5),linewidth=2,
     xlab = "Years since entering moderate frailty state at/after index date",
     ylab = "Probability",
     cex=0.75,
     xlim = c(0,6),
     main="E. From moderate frailty (severe infection)",
     legend = c("Fit","Mild Frailty","Moderate Frailty","Severe Frailty","Death"),
     legend.pos = "top",
     cols = c("#984EA3","#8B8B00","#E41A1C"), # inspired by brewer.pal(5,"Set1"),
     use.ggplot = T)



plot_un_exp_from_moderate <- plot(probtrans_un_exp,from=3,
     ord=c(3:5),linewidth=2,
     xlab = "Years since entering moderate frailty state at/after index date",
     ylab = "Probability",
     cex=0.75,
     xlim = c(0,6),
     main="F. From moderate frailty (no severe infection)",
     legend = c("Fit","Mild Frailty","Moderate Frailty","Severe Frailty","Death"),
     legend.pos = "top",
     cols = c("#984EA3","#8B8B00","#E41A1C"), # inspired by brewer.pal(5,"Set1"),
     use.ggplot = T)

# combine plots
stacked_plot_from_moderate <- ggarrange(plot_exp_from_moderate + ggtitle("E. From moderate frailty (no severe infection)"),
                          plot_un_exp_from_moderate + ggtitle("F. From moderate frailty (no severe infection)"),
                          common.legend = T,
                          nrow=1,hjust = -0)
```

```{r stacked_plots_from_severe}

# From mild
plot_exp_from_severe <- plot(probtrans_exp,from=4,
     ord=c(4:5),linewidth=2,
     xlab = "Years since entering severe frailty state at/after index date",
     ylab = "Probability",
     cex=0.75,
     xlim = c(0,6),
     main="G. From severe frailty (severe infection)",
     legend = c("Fit","Mild Frailty","Moderate Frailty","Severe Frailty","Death"),
     legend.pos = "top",
     cols = c("#8B8B00","#E41A1C"), # inspired by brewer.pal(5,"Set1"),
     use.ggplot = T)


plot_un_exp_from_severe <- plot(probtrans_un_exp,from=4,
     ord=c(4:5),linewidth=2,
     xlab = "Years since entering severe frailty state at/after index date",
     ylab = "Probability",
     cex=0.75,
     xlim = c(0,6),
     main="H. From severe frailty (no severe infection)",
     legend = c("Fit","Mild Frailty","Moderate Frailty","Severe Frailty","Death"),
     legend.pos = "top",
     cols = c("#8B8B00","#E41A1C"), # inspired by brewer.pal(5,"Set1"),
     use.ggplot = T) 

# combine plots
stacked_plot_from_severe <- ggarrange(plot_exp_from_severe + ggtitle("G. From severe frailty (no severe infection)"),
                          plot_un_exp_from_severe + ggtitle("H. From severe frailty (no severe infection)"),
                          common.legend = T,
                          nrow=1,hjust = -0)
```

```{r combine_all_plots}

# combine plots
stacked_plots <- ggarrange(plot_exp_from_fit + ggtitle("A. From fit status (severe infection)"),
                          plot_un_exp_from_fit + ggtitle("B. From fit status (no severe infection)"),
                          plot_exp_from_mild + ggtitle("C. From mild frailty (severe infection)"),
                          plot_un_exp_from_mild + ggtitle("D. From mild frailty (no severe infection)"),
                          plot_exp_from_moderate + ggtitle("E. From moderate frailty (severe infection)"),
                          plot_un_exp_from_moderate + ggtitle("F. From moderate frailty (no severe infection)"),
                          plot_exp_from_severe + ggtitle("G. From severe frailty (severe infection)"),
                          plot_un_exp_from_severe + ggtitle("H. From severe frailty (no severe infection)"),
                          common.legend = T,
                          nrow=4,ncol = 2)
```

```{r mean_state_occupancy_times}

# Expected length of stay in each state; up to 5 years ff-up
# To use the ELOS() in the mstate package
# the time is from study start to end of follow-up
# the ELOS for the absorbing state looks confusing but what it actually means is the expected time since death at study end or the last event time

ELOS(probtrans_exp)
ELOS(probtrans_un_exp)

round( ELOS(probtrans_exp,tau =5),1)
round( ELOS(probtrans_un_exp,tau =5),1)

# exposed
states <- c("Fit","Mild","Moderate","Severe","Death")
 
matrix(ELOS(probtrans_exp,tau =5),nrow = 5,
        dimnames = list(states,states)) |> 
  data.frame() |> rownames_to_column() |>
  mutate(exposed=1,.after=1) |>
  # un_exposed
  rbind(matrix(ELOS(probtrans_un_exp,tau =5),nrow = 5,
        dimnames = list(states,states)) |> 
  data.frame() |> rownames_to_column() |>
    mutate(exposed=0,.after=1)) |> 
  mutate(across(Fit:Death,~round(.x,2))) |>
  gt(rowname_col = "rowname",caption = "Expected length of stay in the starting states before transitioning (5-years follow-up)")
```

## Sensitivity analyses

### Adjust main regression models for ethnicity
```{r all_mstate_cox_ethnicity}
all_mstate_cox_ethnicity <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_ethnicity,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data =  filter(collect(inf_frailty_cprd_mstate_expcovs),
                                               if_all(all_of(c(covariates,ethnicity)),~(!is.na(.x) & .x != "Missing"))))

#
 forestplot_mstate_HRs(model = readRDS("./results/all_mstate_cox_ethnicity.rds"),
                                            events_exposed = events(filter(inf_frailty_cprd_mstate,
                                                                           exposed==1,
                                                                           if_all(all_of(c(covariates,ethnicity)),~(!is.na(.x) & .x != "Missing")))),
                                            events_un_exposed = events(filter(inf_frailty_cprd_mstate,
                                                                           exposed==0,
                                                                           if_all(all_of(c(covariates,ethnicity)),~(!is.na(.x) & .x != "Missing")))),
                       by_sub_group = F, 
                       plot_save_name = "Figure_S3_mstate")
```

### Community acquired infections: redefine severe infections as:

 1. At least one gp recorded infection diagnostic code on the date of hospital admission or within 7 days prior
 2. Any dispensed antimicrobial on the date of hospital admission or within 7 days prior
 3. Both 1 and 2
 
```{r comm_inf_setidss}
#
comm_inf_setids_1 <- collect(inf_frailty_cprd) |>
  mutate(setid = as.numeric(factor(setid))) |>
  group_by(setid) |> 
  filter(any(gp_infection_7dy_lookback == 1 & exposed ==1)) |>
  ungroup() |>
  pull(setid)

#

comm_inf_setids_2 <- collect(inf_frailty_cprd) |>
  mutate(setid = as.numeric(factor(setid))) |>
  group_by(setid) |> 
  filter(any(antibio_7dy_lookbck == 1 & exposed ==1)) |>
  ungroup() |>
  pull(setid)
#

comm_inf_setids_3 <- collect(inf_frailty_cprd) |>
  mutate(setid = as.numeric(factor(setid))) |>
  group_by(setid) |> 
  filter(any(gp_infection_7dy_lookback == 1 & antibio_7dy_lookbck == 1 & exposed ==1)) |>
  ungroup() |>
  pull(setid)

```

```{r community_acquired_infections_1}
all_mstate_cox_comm_inf_1  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),setid %in% comm_inf_setids_1,
                                               if_all(all_of(covariates),~(!is.na(.x) & .x != "Missing"))))
```

```{r community_acquired_infections_2}
all_mstate_cox_comm_inf_2  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),setid %in% comm_inf_setids_2,
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
```

```{r community_acquired_infections_3}
all_mstate_cox_comm_inf_3  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),setid %in% comm_inf_setids_3,
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
```

```{r forest_plot_comm_inf}

plot_df_comm_inf <- rbind( # prior primary care infection diagnosis
  extract_model_data(model = readRDS("./results/all_mstate_cox_comm_inf_1.rds"),
                     events_exposed = events(filter(inf_frailty_cprd_mstate,setid %in% comm_inf_setids_1,exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,setid %in% comm_inf_setids_1,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 1,sub_group_desc="Primary care infection"),
  # prior primary care antibiotic prescription
   extract_model_data(model = readRDS("./results/all_mstate_cox_comm_inf_2.rds"),
                     events_exposed = events(filter(inf_frailty_cprd_mstate,setid %in% comm_inf_setids_2,exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,setid %in% comm_inf_setids_2,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 2,sub_group_desc="Primary care antibiotic"),
  # Both
  extract_model_data(model = readRDS("./results/all_mstate_cox_comm_inf_3.rds"),
                     events_exposed = events(filter(inf_frailty_cprd_mstate,setid %in% comm_inf_setids_3,exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,setid %in% comm_inf_setids_3,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 3,sub_group_desc="Both")) |>
  arrange(term_description,sub_group) |>
  mutate(term_description = if_else(duplicated(term_description),NA_character_,term_description)) 

# Plot
forestplot_mstate_HRs(model = plot_df_comm_inf,
                      height = 40,clip = 10,width = 32,
                      plot_save_name = "Figure_S4_mstate",
                      h_5 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_8 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_11 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_14 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_17 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_20 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_23 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_25 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_28 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2))
```

### Restrict the study to periods with likely more frequent recording of some of the deficits included in the eFI due to changes to the Quality Outcomes Framework (QOF) indicators and other policy changes over time; 

1) 2012 onwards  
2) 2014 onwards 

```{r sensitivity_2012_onwards}

all_mstate_cox_2012  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),indexdate >= as.Date("2012-01-01"),
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
```


```{r sensitivity_2014_onwards}

all_mstate_cox_2014  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),indexdate >= as.Date("2014-01-01"),
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
```

```{r forest_plot_qof_periods}

plot_df_qof_periods <- rbind( # prior primary care infection diagnosis
  extract_model_data(model = all_mstate_cox_2012,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,indexdate >= as.Date("2012-01-01"),exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,indexdate >= as.Date("2012-01-01"),
                                                                           exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 1,sub_group_desc="2012 onwards"),
  # prior primary care antibiotic prescription
   extract_model_data(model = all_mstate_cox_2014,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,indexdate >= as.Date("2014-01-01"),exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,indexdate >= as.Date("2014-01-01"),
                                                                           exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 2,sub_group_desc="2014 onwards")) |>
  arrange(term_description,sub_group) |>
  mutate(term_description = if_else(duplicated(term_description),NA_character_,term_description)) 

# Plot
forestplot_mstate_HRs(model = plot_df_qof_periods,
                      height = 40,clip = 10,width=32,
                      plot_save_name = "Figure_S5_mstate",
                      h_4 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_6 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_8 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_10 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_12 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_14 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_16 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_18 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_20 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2))
```

## Secondary analyses: Effect modification

Fit sub group models to explore effect modification by:

- Age group
- Sex
- Deprivation quintiles
- Diabetes
- Dementia
- Care home residency
- Severe infection (exposure) type (bacterial, viral, fungal, parasitic)
- Severe infection (exposure) site (sepsis,uti,ssti,mening_enceph,lrti,gastro)

### EM Age
```{r em_age_65_70}

all_mstate_cox_age_65_70  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(grep("age",expanded_covariates_infection_history,value = T,invert=T),c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),age_bl_cat4 == "1.65-70",
                                               if_all(all_of(c(grep("age",expanded_covariates_infection_history,value = T,invert=T))),~(!is.na(.x) & .x != "Missing"))))
  
```

```{r em_age_71_80}

all_mstate_cox_age_71_80  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(grep("age",expanded_covariates_infection_history,value = T,invert=T),
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),age_bl_cat4 == "2.71-80",
                                               if_all(all_of(c(grep("age",expanded_covariates_infection_history,value = T,invert=T))),~(!is.na(.x) & .x != "Missing"))))
  
```

```{r em_age_81_90}

all_mstate_cox_age_81_90 <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(grep("age",expanded_covariates_infection_history,value = T,invert=T),
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),age_bl_cat4 == "3.81-90",
                                               if_all(all_of(c(grep("age",expanded_covariates_infection_history,value = T,invert=T))),~(!is.na(.x) & .x != "Missing"))))
  
```

```{r em_age_90}

all_mstate_cox_age_90 <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(grep("age",expanded_covariates_infection_history,value = T,invert=T),
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),age_bl_cat4 == "4.>90",
                                               if_all(all_of(c(grep("age",expanded_covariates_infection_history,value = T,invert=T))),~(!is.na(.x) & .x != "Missing"))))
```

```{r forest_plot_age}

plot_df_age <- rbind( # Age 65_70
  extract_model_data(model = all_mstate_cox_age_65_70,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,age_bl_cat4 == "1.65-70",exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,age_bl_cat4 == "1.65-70",exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 1,sub_group_desc="65-70 Years"),
   # Age 71_80
  extract_model_data(model = all_mstate_cox_age_71_80,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,age_bl_cat4 == "2.71-80",exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,age_bl_cat4 == "2.71-80",exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 2,sub_group_desc="71-80 Years"),
   # Age 81_90
  extract_model_data(model = all_mstate_cox_age_81_90,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,age_bl_cat4 == "3.81-90",exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,age_bl_cat4 == "3.81-90",exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 3,sub_group_desc="81-90 Years"),
   # Age >90
  extract_model_data(model = all_mstate_cox_age_90,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,age_bl_cat4 == "4.>90",exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,age_bl_cat4 == "4.>90",exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 4,sub_group_desc=">90 Years")) |>
  arrange(term_description,sub_group) |>
  mutate(term_description = if_else(duplicated(term_description),NA_character_,term_description)) 

# Plot
forestplot_mstate_HRs(model = plot_df_age,
                      height = 45,clip = 8,width = 32,
                      plot_save_name = "Figure_S6_mstate",
                      h_6 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_10 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_14 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_18 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_22 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_26 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_30 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_34 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_38 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2))
```

### EM sex
```{r em_sex_female}

all_mstate_cox_female  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(grep("gender",expanded_covariates_infection_history,value = T,invert=T),
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),gender == "Female",
                                               if_all(all_of(c(grep("gender",expanded_covariates_infection_history,value = T,invert=T))),~(!is.na(.x) & .x != "Missing"))))
```

```{r em_sex_male}

all_mstate_cox_male  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(grep("gender",expanded_covariates_infection_history,value = T,invert=T),
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),gender == "Male",
                                               if_all(all_of(c(grep("gender",expanded_covariates_infection_history,value = T,invert=T))),~(!is.na(.x) & .x != "Missing"))))
```

```{r forest_plot_sex}

plot_df_sex <- rbind( # prior primary care infection diagnosis
  extract_model_data(model = all_mstate_cox_female,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,gender == "Female", exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,gender == "Female",exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 1,sub_group_desc="Female"),
  # prior primary care antibiotic prescription
   extract_model_data(model = all_mstate_cox_male,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,gender == "Male", exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,gender == "Male",exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 2,sub_group_desc="Male")) |>
  arrange(term_description,sub_group) |>
  mutate(term_description = if_else(duplicated(term_description),NA_character_,term_description)) 

# Plot
forestplot_mstate_HRs(model = plot_df_sex,
                      height = 40,width = 32,
                      plot_save_name = "Figure_S7_mstate",
                      h_4 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_6 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_8 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_10 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_12 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_14 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_16 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_18 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_20 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2))
```

### EM Deprivation
```{r em_dep_1}

all_mstate_cox_dep_1  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(grep("e2019_imd",expanded_covariates_infection_history,value = T,invert=T),
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),e2019_imd == 1,
                                               if_all(all_of(c(grep("e2019_imd",expanded_covariates_infection_history,value = T,invert=T))),~(!is.na(.x) & .x != "Missing"))))
  
```

```{r em_dep_2}
all_mstate_cox_dep_2  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(grep("e2019_imd",expanded_covariates_infection_history,value = T,invert=T),
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),e2019_imd == 2,
                                               if_all(all_of(c(grep("e2019_imd",expanded_covariates_infection_history,value = T,invert=T))),~(!is.na(.x) & .x != "Missing"))))
```

```{r em_dep_3}
all_mstate_cox_dep_3  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(grep("e2019_imd",expanded_covariates_infection_history,value = T,invert=T),
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),e2019_imd == 3,
                                               if_all(all_of(c(grep("e2019_imd",expanded_covariates_infection_history,value = T,invert=T))),~(!is.na(.x) & .x != "Missing"))))
```

```{r em_dep_4}
all_mstate_cox_dep_4  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(grep("e2019_imd",expanded_covariates_infection_history,value = T,invert=T),
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),e2019_imd == 4,
                                               if_all(all_of(c(grep("e2019_imd",expanded_covariates_infection_history,value = T,invert=T))),~(!is.na(.x) & .x != "Missing"))))
```

```{r em_dep_5}
all_mstate_cox_dep_5  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(grep("e2019_imd",expanded_covariates_infection_history,value = T,invert=T),
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),e2019_imd == 5,
                                               if_all(all_of(c(grep("e2019_imd",expanded_covariates_infection_history,value = T,invert=T))),~(!is.na(.x) & .x != "Missing"))))
```

```{r combine_em_dep}

plot_df_deprivation <- rbind( # IMD 1
  extract_model_data(model = all_mstate_cox_dep_1.rds,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,e2019_imd == 1,exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,e2019_imd == 1,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 1,sub_group_desc="IMD 1"),
   # IMD 2
  extract_model_data(model = all_mstate_cox_dep_2.rds,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,e2019_imd == 2,exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,e2019_imd == 2,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 2,sub_group_desc="IMD 2"),
   # IMD 3
  extract_model_data(model = all_mstate_cox_dep_3.rds,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,e2019_imd == 3,exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,e2019_imd == 3,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 3,sub_group_desc="IMD 3"),
   # IMD 4
  extract_model_data(model = all_mstate_cox_dep_4.rds,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,e2019_imd == 4,exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,e2019_imd == 4,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 4,sub_group_desc="IMD 4"),
# IMD 5
extract_model_data(model = all_mstate_cox_dep_5.rds,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,e2019_imd == 5,exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,e2019_imd == 5,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 5,sub_group_desc="IMD 5")) |>
  arrange(term_description,sub_group) |>
  mutate(term_description = if_else(duplicated(term_description),NA_character_,term_description)) 

# Plot
forestplot_mstate_HRs(model = plot_df_deprivation,
                      height = 60,legend_dist = 0.13,width = 32,
                      plot_save_name = "Figure_S8_mstate",
                      h_7 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_12 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_17 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_22 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_27 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_32 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_37 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_42 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_47 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2))
```

### EM Care home residency
```{r em_care_home_no}
all_mstate_cox_care_home_no  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),care_home_residency == 0,
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
```

```{r em_care_home_yes}
all_mstate_cox_care_home_yes  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),care_home_residency == 1,
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
```

```{r combine_em_care_home}
plot_df_care_home <- rbind( # all_mstate_cox_care_home_no
  extract_model_data(model = all_mstate_cox_care_home_no,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,care_home_residency == 0, exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,care_home_residency == 0,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 1,sub_group_desc="Care home: No"),
  # all_mstate_cox_care_home_yes
   extract_model_data(model = all_mstate_cox_care_home_yes),
                     events_exposed = events(filter(inf_frailty_cprd_mstate,care_home_residency == 1, exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,care_home_residency == 1,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 2,sub_group_desc="Care home: Yes")) |>
  arrange(term_description,sub_group) |>
  mutate(term_description = if_else(duplicated(term_description),NA_character_,term_description)) 

# Plot
forestplot_mstate_HRs(model = plot_df_care_home,
                      height = 40,width = 32,
                      plot_save_name = "Figure_S9_mstate",
                      h_4 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_6 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_8 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_10 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_12 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_14 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_16 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_18 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_20 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2))
```

### EM Dementia
```{r em_dementia_no}
all_mstate_cox_dementia_no  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),dementia == 0,
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))

```

```{r em_dementia_yes}
all_mstate_cox_dementia_yes  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),dementia == 1,
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
  
```

```{r combine_em_dementia}
plot_df_dementia <- rbind( # prior primary care infection diagnosis
  extract_model_data(model = all_mstate_cox_dementia_no,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,dementia == 0, exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,dementia == 0,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 1,sub_group_desc="Dementia: No"),
  # prior primary care antibiotic prescription
   extract_model_data(model = all_mstate_cox_dementia_yes,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,dementia == 1, exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,dementia == 1,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 2,sub_group_desc="Dementia: Yes")) |>
  arrange(term_description,sub_group) |>
  mutate(term_description = if_else(duplicated(term_description),NA_character_,term_description)) 

# Plot
forestplot_mstate_HRs(model = plot_df_dementia,
                      height = 40,width = 32,
                      plot_save_name = "Figure_S10_mstate",
                      h_4 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_6 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_8 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_10 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_12 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_14 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_16 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_18 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_20 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2))
```

### EM Diabetes
```{r em_diabetes_no}
all_mstate_cox_diabetes_no  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),diabetes == 0,
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
```

```{r em_diabetes_yes}
all_mstate_cox_diabetes_yes  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,
                                                                    c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),diabetes == 1,
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
```

```{r combine_em_diabetes}
plot_df_diabetes <- rbind( # prior primary care infection diagnosis
  extract_model_data(model = all_mstate_cox_diabetes_no,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,diabetes == 0, exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,diabetes == 0,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 1,sub_group_desc="Diabetes: No"),
  # prior primary care antibiotic prescription
   extract_model_data(model = all_mstate_cox_diabetes_yes,
                     events_exposed = events(filter(inf_frailty_cprd_mstate,diabetes == 1, exposed==1,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     events_un_exposed = events(filter(inf_frailty_cprd_mstate,diabetes == 1,exposed==0,
                                                                           if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 2,sub_group_desc="Diabetes: Yes")) |>
  arrange(term_description,sub_group) |>
  mutate(term_description = if_else(duplicated(term_description),NA_character_,term_description)) 

# Plot
forestplot_mstate_HRs(model = plot_df_diabetes,
                      height = 40,width = 32,
                      plot_save_name = "Figure_S11_mstate",
                      h_4 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_6 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_8 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_10 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_12 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_14 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_16 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_18 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_20 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2))
```


### EM infection pathogen 
```{r em_bacterial}
all_mstate_cox_bacterial  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),inf_type_of_match_grp == "bacterial",
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))

```

```{r em_viral}
all_mstate_cox_viral  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),inf_type_of_match_grp == "viral",
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
```

```{r em_fungal}
all_mstate_cox_fungal  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),inf_type_of_match_grp == "fungal",
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
```

```{r em_parasitic}
all_mstate_cox_parasitic  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),inf_type_of_match_grp == "parasitic",
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
```

```{r combine_em_pathogen}
plot_df_inf_pathogen <- rbind( # inf_pathogen 65_70
  extract_model_data( model = all_mstate_cox_bacterial,
  events_exposed = events(filter(inf_frailty_cprd_mstate,inf_type_of_match_grp == "bacterial",exposed==1,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
  events_un_exposed = events(filter(inf_frailty_cprd_mstate,inf_type_of_match_grp == "bacterial",exposed==0,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 1,sub_group_desc="Bacterial"),
   # inf_pathogen 71_80
  extract_model_data( model = all_mstate_cox_viral,
  events_exposed = events(filter(inf_frailty_cprd_mstate,inf_type_of_match_grp == "viral",exposed==1,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
  events_un_exposed = events(filter(inf_frailty_cprd_mstate,inf_type_of_match_grp == "viral",exposed==0,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 2,sub_group_desc="Viral"),
   # inf_pathogen 81_90
  extract_model_data( model = all_mstate_cox_fungal,
  events_exposed = events(filter(inf_frailty_cprd_mstate,inf_type_of_match_grp == "fungal",exposed==1,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
  events_un_exposed = events(filter(inf_frailty_cprd_mstate,inf_type_of_match_grp == "fungal",exposed==0,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 3,sub_group_desc="Fungal"),
   # inf_pathogen >90
  extract_model_data( model = all_mstate_cox_parasitic,
  events_exposed = events(filter(inf_frailty_cprd_mstate,inf_type_of_match_grp == "parasitic",exposed==1,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
  events_un_exposed = events(filter(inf_frailty_cprd_mstate,inf_type_of_match_grp == "parasitic",exposed==0,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                     sub_group_level = 4,sub_group_desc="Parasitic")) |>
  arrange(term_description,sub_group) |>
  mutate(term_description = if_else(duplicated(term_description),NA_character_,term_description)) 

# Plot
forestplot_mstate_HRs(model = plot_df_inf_pathogen,
                      height = 45,clip=10,width = 32,
                      plot_save_name = "Figure_S12_mstate",
                      h_6 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_10 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_14 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_18 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_22 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_26 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_30 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_34 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2),
                      h_38 = gpar(col = "black", columns = 1:5, lty = "solid",lwd=2))
```

### EM infection type

```{r em_sepsis}
all_mstate_cox_sepsis  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),inf_site_of_match_grp == "sepsis",
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
  
#
forestplot_mstate_HRs(model = all_mstate_cox_sepsis,
                                            events_exposed = events(filter(inf_frailty_cprd_mstate,inf_site_of_match_grp == "sepsis",exposed==1,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                                            events_un_exposed = events(filter(inf_frailty_cprd_mstate,inf_site_of_match_grp == "sepsis",exposed==0,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                                            plot_save_name = "forest_plot_sepsis")
```

```{r em_uti}
all_mstate_cox_uti  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),inf_site_of_match_grp == "uti",
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))

  
#
forestplot_mstate_HRs(model = all_mstate_cox_uti,
                                            events_exposed = events(filter(inf_frailty_cprd_mstate,inf_site_of_match_grp == "uti",exposed==1,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                                            events_un_exposed = events(filter(inf_frailty_cprd_mstate,inf_site_of_match_grp == "uti",exposed==0,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                                            plot_save_name = "forest_plot_uti")
```

```{r em_ssti}
all_mstate_cox_ssti  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),inf_site_of_match_grp == "ssti",
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
  
#
forestplot_mstate_HRs(model = all_mstate_cox_ssto,
                                            events_exposed = events(filter(inf_frailty_cprd_mstate,inf_site_of_match_grp == "ssti",exposed==1,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                                            events_un_exposed = events(filter(inf_frailty_cprd_mstate,inf_site_of_match_grp == "ssti",exposed==0,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                                            plot_save_name = "forest_plot_ssti")
```

```{r em_lrti}
all_mstate_cox_lrti  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),inf_site_of_match_grp == "lrti",
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
  
#
forestplot_mstate_HRs(model = all_mstate_cox_lrti,
                                            events_exposed = events(filter(inf_frailty_cprd_mstate,inf_site_of_match_grp == "lrti",exposed==1,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                                            events_un_exposed = events(filter(inf_frailty_cprd_mstate,inf_site_of_match_grp == "lrti",exposed==0,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                                            plot_save_name = "forest_plot_lrti")
```

```{r em_mening_enceph}
all_mstate_cox_mening_enceph  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                  c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                 method = "breslow",
                                 data = filter(collect(inf_frailty_cprd_mstate_expcovs),inf_site_of_match_grp == "mening_enceph",
                                               if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))
  
#
forestplot_mstate_HRs(model = all_mstate_cox_mening_enceph,
                                            events_exposed = events(filter(inf_frailty_cprd_mstate,inf_site_of_match_grp == "mening_enceph",exposed==1,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                                            events_un_exposed = events(filter(inf_frailty_cprd_mstate,inf_site_of_match_grp == "mening_enceph",exposed==0,
                                                                           if_all(all_of(c(covariates)),
                                                                                  ~(!is.na(.x) & .x != "Missing")))),
                                            plot_save_name = "forest_plot_mening_enceph")
```

# For gastro:
**Warning message:In grep(c("age", "harmful_alcohol_use"), expanded_covariates_infection_history) :argument 'pattern' has length > 1 and only the first element will be used**

# There is perfect success prediction for age_bl_cat43.6 (81-90) harmful_alcohol_use.6 in transition 6
# So to remove variables only for that transition

  
```{r em_gastro}
all_mstate_cox_gastro  <- coxph(as.formula(paste("Surv(Tstart,Tstop,status)","~",
                                                        c(glue_collapse(c(expanded_covariates_infection_history,c("strata(trans)","cluster(setid)")),sep = "+")))),
                                       method = "breslow",
                                       data = filter(collect(inf_frailty_cprd_mstate_expcovs),inf_site_of_match_grp == "gastro",
                                                     if_all(all_of(c(covariates)),~(!is.na(.x) & .x != "Missing"))))

#
forestplot_mstate_HRs(model = all_mstate_cox_gastro,
                      events_exposed = events(filter(inf_frailty_cprd_mstate,inf_site_of_match_grp == "gastro",exposed==1,
                                                     if_all(all_of(c(covariates)),
                                                            ~(!is.na(.x) & .x != "Missing")))),
                      events_un_exposed = events(filter(inf_frailty_cprd_mstate,inf_site_of_match_grp == "gastro",exposed==0,
                                                        if_all(all_of(c(covariates)),
                                                               ~(!is.na(.x) & .x != "Missing")))),
                      plot_save_name = "forest_plot_gastro")

```
