---
title: "Create tidy data ready for analysis of study on Infections and Frailty in CPRD-UK"
format: html
author: Kwabena Asare
toc: true
toc-location: left
number-sections: true
date: "`r lubridate::today()`"
tbl-cap-location: top
editor_options: 
  chunk_output_type: console
execute: 
  include: true
  echo: false
editor: 
  markdown: 
    wrap: 72
---

## Function to define covariates/variables with codelist

# objects needed
1. A codelist that defines the variable. 
  
  -The codelist should have 2 main columns; 
  - The code; ICD, SNOWMED etc
  - The description; 
  
2. A dataset of observations/individuals; **This is the observation file**
  the data set should have 2 main variables;

  - The patid; unique identifier of each observation
  - The medical code; all medical codes that represent any observation such as diagnosis or race/ethnicity in the EHR

3. Approach. for each codelist, join the codelist and obsertion file using the medical code as join key
  - then the description column from the codelist will now be the new variable.
  
```{r function_to_def_covariate}

# create a function
# Arguments of the function
# 1. data_files = list of files that contain the data sets;these are the observation files from CPRD
# 2. codelist = a list of codes that define a particular medical observation such as diagnosis
# codelist has code and description
# 3. days_look_back = What range of look back in time should be used to select the cohort_entry observation
# Ever look back for most variables based on study protocol; 
# other default variables
  # medcodeid = the code
  # term= description of code 

def_covariate <- function(data_files = observations,codelist) {
  
  data <- map(data_files, .progress = TRUE,function(x) {
    #
      collect(x) |>
      fselect(patid,medcodeid,obsdate) |>
      fmutate(obsdate = as.Date(obsdate,format="%d/%m/%Y")) |>
      join(select(collect(codelist),medcodeid,term),
       how = "inner",
       on="medcodeid") |> 
      fselect(patid,term,obsdate)}) |> reduce(rowbind)
  # 
  data |>
    funique(cols = c("patid","obsdate")) |>
    join(fselect(collect(cohort_matched_2), patid,setid,indexdate,enddate),
         how = "left",
         on = "patid",
         multiple =  TRUE) |> # patids appearing as both exposed and unexposed
    # below is important for patids appearing as both exposed and un-exposed but with diff indexdate
    # filter observations from the lookback to the indexdate
    fsubset(obsdate <= indexdate) |>
    funique(cols=c("patid","setid","obsdate")) |>
    fselect(patid,setid,obsdate,term) |>
    group_by(patid,setid) |>
    arrange(obsdate,.by_group = T) |>
    slice_tail() |> # to take the most recent obsdate
    ungroup() |>
    fselect(patid,setid,term)
}
```

## Exposure: severe infections type and site 

```{r exposure_tidy}
# create infection type category
# create infection site category

#
  collect(cohort_matched_2) |>
  fselect(patid,setid,ICD,exposed) |>
  join(distinct(mutate(infection_codelist, ICD=str_sub(`ICD-10 code`,1,5)),ICD,.keep_all = T),on ="ICD",how="left") |> 
  # infection type
  fmutate(inf_type = case_when(bacterial %in% "1" ~ "bacterial",
                              viral %in% "1" ~ "viral",
                              fungal %in% "1" ~ "fungal",
                              parasitic %in% "1" ~ "parasitic")) |>
  # infection site
  fmutate(inf_site = case_when(sepsis %in% 1 ~ "sepsis",
                              uti %in% 1 ~ "uti",
                              ssti %in% 1 ~ "ssti",
                              mening_enceph %in% 1 ~ 'mening_enceph',
                              lrti %in% 1 ~ 'lrti',
                              gastro %in% 1 ~ 'gastro')) |>
  # infection type and site of setid; matchedd group
  group_by(setid) |>
  arrange(exposed,.by_group = T) |> # this puts exposed = 1 as the last obs among same setid
  fmutate(inf_type_of_match_grp = last(inf_type),.after = inf_type) |>
  fmutate(inf_site_of_match_grp = last(inf_site),.after = inf_site) |>
  ungroup() |>
  fselect(patid,setid,inf_type_of_match_grp,inf_site_of_match_grp) 
```

## Covariates defined with codes  

```{r ethnicity_tidy}
#
ethnicity_data <- def_covariate(data_files = observations,codelist = ethnicity_codes) |>
  to_duckdb(con=my_con_integer64)
#
collect(ethnicity_data) |>
  join(select(collect(cohort_matched_2),patid,setid),on=c("patid","setid"),how="full") |>
  fmutate(ethnicity = if_else(term %in% c("5. Not Stated","") | is.na(term),
                                  "Missing",term)) 
```

```{r alcohol_use_tidy}
#
  alcohol_data <- def_covariate(data_files = observations,codelist = alcohol_codes) |>
  to_duckdb(con=my_con_integer64)
#
  collect(alcohol_data) |>
  join(select(collect(cohort_matched_2),patid,setid),on=c("patid","setid"),how="full") |>
  fmutate(harmful_alcohol_use = if_else(!is.na(term),1,0)) 
```

```{r smoking_tidy}
#
 smoking_data <- def_covariate(data_files = observations,codelist = smoking_codes) |>
  to_duckdb(con=my_con_integer64)

#
  collect(smoking_data) |>
  fmutate(smoking = case_when(str_detect(term,"(ex|EX|Ex|eX|[Ss]topped|ceased).*[Ss]mok") ~"2.Ex-smoker",
                                       str_detect(term,"([Nn]ever|[Nn]on).*[sS]mok") ~"1.Non-smoker",
                                       .default = "3.Current smoker")) |>
  join(select(collect(cohort_matched_2),patid,setid),on=c("patid","setid"),how="full") |>
  fmutate(smoking =if_else(is.na(smoking),"Missing",smoking)) 
```

```{r dementia_tidy}
#  
dementia_data <- def_covariate(data_files = observations,codelist = dementia_codes)  |>
  to_duckdb(con=my_con_integer64)
#
  collect(dementia_data) |>
  join(select(collect(cohort_matched_2),patid,setid),on=c("patid","setid"),how="full") |>
  fmutate(dementia = if_else(!is.na(term),1,0)) 
```

```{r diabetes_tidy}
# 
 diabetes_data <- def_covariate(data_files = observations,codelist = diabetes_codes) |>
  to_duckdb(con=my_con_integer64)
#
  collect(diabetes_data) |>
  join(select(collect(cohort_matched_2),patid,setid),on=c("patid","setid"),how="full") |>
  fmutate(diabetes = if_else(!is.na(term),1,0)) 
```

```{r care_home_residency_tidy}
#
 care_home_residency_data <- def_covariate(data_files = observations,codelist = care_home_residency_codes) |>
  to_duckdb(con=my_con_integer64)
#
collect(care_home_residency_data) |>
  join(select(collect(cohort_matched_2),patid,setid),on=c("patid","setid"),how="full") |>
  fmutate(care_home_residency = if_else(!is.na(term),1,0)) 
```

```{r demographics_tidy}

# create follow-up duration variable
# aka survival time in days
  cohort_matched_2 |> 
  collect() |>
  left_join(transmute(practice_imd,pracid,e2019_imd_5_prac=e2019_imd_5),join_by(pracid)) |>
  left_join(transmute(patient_2019_imd,patid,e2019_imd_5_pat=e2019_imd_5),join_by(patid )) |>
  fmutate(person_time_days = as.numeric(enddate-indexdate),
         person_time_years = as.numeric(enddate-indexdate)/365) |>
  fmutate(censored_yr1 = if_else(person_time_days > 0, 0,1),
         censored_yr2 = if_else(person_time_days > 365, 0,1),
         censored_yr3 = if_else(person_time_days > 365*2, 0,1),
         censored_yr4 = if_else(person_time_days > 365*3, 0,1),
         censored_yr5 = if_else(person_time_days > 365*4, 0,1)) |>
  # break down censoring into death, lcd/regenddate, study_end_date
  # create only 2 levels, death or others since dates can coincide
    fmutate(follow_up_yr1 = case_when(censored_yr1 == 1 & enddate == cprd_ddate ~ "Died",
                                      censored_yr1 == 1 & (is.na(cprd_ddate) | enddate != cprd_ddate)  ~ "lcd_regenddate_studyend",
                                      .default = "Active"),
         follow_up_yr2 = case_when(censored_yr2 == 1 & enddate== cprd_ddate ~ "Died",
                                      censored_yr2 == 1 & (is.na(cprd_ddate) | enddate != cprd_ddate)   ~ "lcd_regenddate_studyend",
                                      .default = "Active"),
         follow_up_yr3 = case_when(censored_yr3 == 1 & enddate== cprd_ddate ~ "Died",
                                      censored_yr3 == 1 & (is.na(cprd_ddate) | enddate != cprd_ddate)   ~ "lcd_regenddate_studyend",
                                      .default = "Active"),
         follow_up_yr4 = case_when(censored_yr4 == 1 & enddate== cprd_ddate ~ "Died",
                                      censored_yr4 == 1 & (is.na(cprd_ddate) | enddate != cprd_ddate)   ~ "lcd_regenddate_studyend",
                                      .default = "Active"),
         follow_up_yr5 = case_when(censored_yr5 == 1 & enddate== cprd_ddate ~ "Died",
                                      censored_yr5 == 1 & (is.na(cprd_ddate) | enddate != cprd_ddate)   ~ "lcd_regenddate_studyend",
                                      .default = "Active")) |>
  # Ag3
  fmutate(age_bl = (year(indexdate)-yob),
         ageout = (year(enddate)-yob)) |>
  fmutate(age_bl_cat4 = cut(age_bl,breaks=c(65,70,80,90,Inf),include.lowest=T,
                     labels=c("1.65-70","2.71-80","3.81-90","4.>90")),
         ageout_cat4 = cut(ageout,breaks=c(65,70,80,90,Inf),include.lowest=T,
                     labels=c("1.65-70","2.71-80","3.81-90","4.>90"))) |>
  # Gender
 fmutate(gender = case_when(gender ==1 ~ "Male",gender==2 ~ "Female")) |>
 # IMD
 # if e2019_imd_5_pat available use it otherwise take e2019_imd_5_prac 
 # e2019_imd_5_pat  = patient level IMD
 # e2019_imd_5_prac = practice level IMD
  fmutate(e2019_imd = coalesce(e2019_imd_5_pat,e2019_imd_5_prac),.after = e2019_imd_5_prac) |>
  fmutate(e2019_imd_cat = case_when(e2019_imd ==1 ~"1, least deprived",
                                   e2019_imd ==2~"2, 2nd quintile",
                                   e2019_imd ==3~"3, 3rd quintile",
                                   e2019_imd ==4~"4, 4th quintile",
                                   e2019_imd ==5~"5, Most deprived")) |>
  select(patid,setid,person_time_days,person_time_years,matches("censored|follow_up"),
          age_bl,ageout,age_bl_cat4,ageout_cat4,gender,e2019_imd,e2019_imd_cat) 
```

```{r lookback_gp_infections_tidy}

 map(observations,.progress = TRUE,function(x) {
    #
      collect(x) |>
      fselect(patid,medcodeid,obsdate) |>
      fmutate(obsdate = as.Date(obsdate,format="%d/%m/%Y")) |>
      join(collect(gp_infection_codes) |> select(medcodeid),
       how = "inner",
       on="medcodeid") |> 
      fselect(patid,obsdate)}) |> reduce(rowbind) |>
    funique(cols = c("patid","obsdate")) |>
    join(fselect(collect(cohort_matched_2), patid,setid,indexdate,enddate),
         how = "left",
         on = "patid",
         multiple =  TRUE) |> 
    fsubset(obsdate>indexdate-1835 & obsdate <= indexdate) |>
    funique(cols=c("patid","setid","obsdate")) |>
    group_by(patid,setid) |>
    fmutate(gp_infection_5yr_lookback = case_when(any(obsdate > indexdate-1835 & obsdate <= indexdate) ~ 1),
            gp_infection_7dy_lookback = case_when(any(obsdate > indexdate-7 & obsdate <= indexdate) ~ 1)) |>
    ungroup() |>
    funique(cols=c("patid","setid")) |>
    fselect(patid,setid,gp_infection_5yr_lookback,gp_infection_7dy_lookback) |>
    join(select(collect(cohort_matched_2),patid,setid),on=c("patid","setid"),how="full") |>
    fmutate(gp_infection_5yr_lookback = if_else(!is.na(gp_infection_5yr_lookback),1,0),
         gp_infection_7dy_lookback = if_else(!is.na(gp_infection_7dy_lookback),1,0)) 

```

CPRD antibiotic definition

```{r antibio_7dy_lookbck}

  map(drugissues,.progress = T,function(x) {
      collect(x) |>
      fselect(patid,issuedate,prodcodeid) |>
      fmutate(issuedate=as.Date(issuedate,"%d/%m/%Y")) |>
      join(collect(aurum_product_dictionary),on=c("prodcodeid"),how="inner") |> 
      fsubset(ProductName %in% antimicrobials) |>
      join(select(collect(cohort_matched_2),patid,setid,indexdate,enddate),
       on=c("patid"),
       how="left",
       multiple = TRUE) |>
      fsubset(issuedate > indexdate-7 & issuedate <= indexdate)}) |>
  reduce(rowbind) |>
  funique(cols = c("patid","issuedate")) |>
    # create antibio_7dy_lookbck; outpatient antibiotic prescription 7 days before cohort entry
  fmutate(antibio_7dy_lookbck = 1) |>
  # set missing antibio_7dy_lookbck to 0
  join(select(collect(cohort_matched_2),patid,setid),
       on=c("patid","setid"),
       how="full",
       multiple = TRUE) |>
  fmutate(antibio_7dy_lookbck = if_else(antibio_7dy_lookbck==1,1,0,missing=0)) |>
  funique(cols = c("patid","setid")) |>
  fselect(patid,setid,antibio_7dy_lookbck) 
```

```{r consultations_inyear_before_cohort_entry}
 map(consultations,.progress = T,function(x) {
  collect(x)|> 
  fselect(patid,consdate) |>
  fmutate(consdate = as.Date(consdate,format="%d/%m/%Y")) |>
     # allow 1 yr before indexdate
     # add obs with no consultation dates; leave consultations == 0
     join(fselect(collect(cohort_matched_2),patid,setid,indexdate),
         on ="patid",
         how="left",
         multiple = TRUE) |>
     fsubset(consdate > indexdate-365 & consdate < indexdate) |>
     fmutate(one_consult_1y_lookbck = if_else(consdate > indexdate-365 & consdate < indexdate,1,0)) |>
     fselect(patid,setid,one_consult_1y_lookbck) |>
     group_by(patid,setid) |>
     arrange(one_consult_1y_lookbck,.by_group = T) |>
     slice(1) |>
     ungroup()
     }) |> 
  reduce(rowbind) |>
  # join full and set missing one_consult_1y_lookbck to NA
  join(fselect(collect(cohort_matched_2),patid,setid),
         on =c("patid","setid"),
         how="full",
         multiple = TRUE) |>
  fmutate(one_consult_1y_lookbck=if_else(is.na(one_consult_1y_lookbck),0,one_consult_1y_lookbck)) 
```

## Outcome: eFI data
## Frailty data 1: Create eFI deficits at baseline/indexdate and at years 1, 2, 3, 4, and 5

Approach:
1. Start with observation file
2. Join observation file with codelist on medcodeid,
3. If code matches, deficit = 1, otherwise deficit = 0
4. If patid censored, deficit = NA otherwise bullet 3
5. First define eFI at index date and update at each ff-up if a new deficit is recorded, 
otherwise maintain previous eFI
6. 36 deficits in total, 35 are are commordities/health etc defined with codelist and 
poypharmacy defined by counting number of prescribed medicines in BNF chapters 1-15 

### 35 eFI deficits

```{r 35_eFI_deficits}

# 35 eFI deficits data 
# Objects needed

# file paths to codelist
# use list.files() function to get the list of file paths
eFI_deficits_codelists_file_paths

# a vector of the names of deficits
deficits_names <- map_chr(eFI_deficits_codelists_file_paths, ~str_extract(.x,pattern = "(.*/)(.*)( .* codelist.csv$)",group = 2))
 
# a named list of the codelists
deficits_code_list <- setNames(map(eFI_deficits_codelists_file_paths, ~unique(pull(collect(open_csv_dataset(.x)),medcodeid))),
                              deficits_names)
# all codelits combined: vector
deficits_code_df <- map(deficits_code_list,~as.data.frame(.x)) |> reduce(rowbind) |>
  rename_with(~"medcodeid")

# define deficit based on codelist
# iterate the ff code on the observation files and combine result into a data set of the 35 deficits

  map(observations,.progress = TRUE,function(x) {
  
  collect(x) |> 
    fselect(patid,obsdate,medcodeid) |>
    funique(cols=c("patid","obsdate","medcodeid")) |>
    fmutate(obsdate = as.Date(obsdate,format="%d/%m/%Y")) |>
    fsubset(medcodeid %in% deficits_code_df$medcodeid) |>
    fmutate(deficit_name=case_when(medcodeid %in% deficits_code_list[[1]] ~ deficits_names[[1]],
                              medcodeid %in% deficits_code_list[[2]] ~ deficits_names[[2]],
                              medcodeid %in% deficits_code_list[[3]] ~ deficits_names[[3]],
                              medcodeid %in% deficits_code_list[[4]] ~ deficits_names[[4]],
                              medcodeid %in% deficits_code_list[[5]] ~ deficits_names[[5]],
                              medcodeid %in% deficits_code_list[[6]] ~ deficits_names[[6]],
                              medcodeid %in% deficits_code_list[[7]] ~ deficits_names[[7]],
                              medcodeid %in% deficits_code_list[[8]] ~ deficits_names[[8]],
                              medcodeid %in% deficits_code_list[[9]] ~ deficits_names[[9]],
                              medcodeid %in% deficits_code_list[[10]] ~ deficits_names[[10]],
                              medcodeid %in% deficits_code_list[[11]] ~ deficits_names[[11]],
                              medcodeid %in% deficits_code_list[[12]] ~ deficits_names[[12]],
                              medcodeid %in% deficits_code_list[[13]] ~ deficits_names[[13]],
                              medcodeid %in% deficits_code_list[[14]] ~ deficits_names[[14]],
                              medcodeid %in% deficits_code_list[[15]] ~ deficits_names[[15]],
                              medcodeid %in% deficits_code_list[[16]] ~ deficits_names[[16]],
                              medcodeid %in% deficits_code_list[[17]] ~ deficits_names[[17]],
                              medcodeid %in% deficits_code_list[[18]] ~ deficits_names[[18]],
                              medcodeid %in% deficits_code_list[[19]] ~ deficits_names[[19]],
                              medcodeid %in% deficits_code_list[[20]] ~ deficits_names[[20]],
                              medcodeid %in% deficits_code_list[[21]] ~ deficits_names[[21]],
                              medcodeid %in% deficits_code_list[[22]] ~ deficits_names[[22]],
                              medcodeid %in% deficits_code_list[[23]] ~ deficits_names[[23]],
                              medcodeid %in% deficits_code_list[[24]] ~ deficits_names[[24]],
                              medcodeid %in% deficits_code_list[[25]] ~ deficits_names[[25]],
                              medcodeid %in% deficits_code_list[[26]] ~ deficits_names[[26]],
                              medcodeid %in% deficits_code_list[[27]] ~ deficits_names[[27]],
                              medcodeid %in% deficits_code_list[[28]] ~ deficits_names[[28]],
                              medcodeid %in% deficits_code_list[[29]] ~ deficits_names[[29]],
                              medcodeid %in% deficits_code_list[[30]] ~ deficits_names[[30]],
                              medcodeid %in% deficits_code_list[[31]] ~ deficits_names[[31]],
                              medcodeid %in% deficits_code_list[[32]] ~ deficits_names[[32]],
                              medcodeid %in% deficits_code_list[[33]] ~ deficits_names[[33]],
                              medcodeid %in% deficits_code_list[[34]] ~ deficits_names[[34]],
                              medcodeid %in% deficits_code_list[[35]] ~ deficits_names[[35]])) |>
    fselect(patid,obsdate,deficit_name) |>
    fmutate(deficit_yes_no = case_when(!is.na(deficit_name) ~ 1)) |>
    join(fselect(collect(cohort_matched_2),patid,setid,indexdate,enddate),
         on = "patid",
         how ="left",
         multiple = TRUE) |> # patids appearing as both exposed and unexposed
    fsubset(obsdate > as.Date(-Inf) & obsdate <= indexdate+1825) |>
    fmutate(time_point = case_when(obsdate > as.Date(-Inf) & obsdate <= indexdate-14 ~ "bl",
                                   obsdate > indexdate & obsdate <= indexdate+365 ~ "yr1",
                                   obsdate > indexdate+365 & obsdate <= indexdate+730 ~ "yr2",
                                   obsdate > indexdate+730 & obsdate <= indexdate+1095 ~ "yr3",
                                   obsdate > indexdate+1095 & obsdate <= indexdate+1460 ~ "yr4",
                                   obsdate > indexdate+1460 & obsdate <= indexdate+1825 ~ "yr5")) |>
    fsubset(!is.na(time_point)) |>
    funique(cols=c("patid","setid","deficit_name","time_point")) |>
    fselect(patid,setid,deficit_yes_no,deficit_name,time_point) |>
    pivot_wider(names_from = c(deficit_name,time_point),
                values_from = deficit_yes_no,
                names_glue = "{deficit_name}_{time_point}",
                values_fill = NA) 
  }) |> # the result of this is a list
  reduce(rowbind,fill=T) |> 
  # some observations will not match any deficit; add them 
  join(fselect(collect(cohort_matched_2),patid,setid,indexdate,enddate),
         on = c("patid","setid"),
         how ="full") |>
  fmutate(person_time_days = as.numeric(enddate - indexdate)) |>
  # create deficits: approach -1
  # first create eFI at cohort entry (1 or 0)
  # then if new deficit recorded at ff-up, 1, otherwise, previous/lag deficit
  mutate(across(matches(glue("{deficits_names}")),~ if_else(is.na(.x),0,.x))) |> # all deficits over time
  mutate(across(matches(glue("{deficits_names}_yr1")),~ if_else(.x==1,.x,get(glue("{str_replace(cur_column(),'yr1','bl')}"))))) |>
  mutate(across(matches(glue("{deficits_names}_yr2")),~ if_else(.x==1,.x,get(glue("{str_replace(cur_column(),'yr2','yr1')}"))))) |>
  mutate(across(matches(glue("{deficits_names}_yr3")),~ if_else(.x==1,.x,get(glue("{str_replace(cur_column(),'yr3','yr2')}"))))) |>
  mutate(across(matches(glue("{deficits_names}_yr4")),~ if_else(.x==1,.x,get(glue("{str_replace(cur_column(),'yr4','yr3')}"))))) |>
  mutate(across(matches(glue("{deficits_names}_yr5")),~ if_else(.x==1,.x,get(glue("{str_replace(cur_column(),'yr5','yr4')}"))))) |>
  # If not censored and no match for deficit code, deficit = 0
  # If censored, deficit = NA
  mutate(across(matches(glue("{deficits_names}_yr1")),~ if_else(person_time_days > 0,.x,NA_integer_))) |>
  mutate(across(matches(glue("{deficits_names}_yr2")),~ if_else(person_time_days > 365,.x,NA_integer_))) |>
  mutate(across(matches(glue("{deficits_names}_yr3")),~ if_else(person_time_days > 730,.x,NA_integer_)))|>
  mutate(across(matches(glue("{deficits_names}_yr4")),~ if_else(person_time_days > 1095,.x,NA_integer_)))|>
  mutate(across(matches(glue("{deficits_names}_yr5")),~ if_else(person_time_days > 1460,.x,NA_integer_))) 
   
  #
  eFI_35_deficits_data |>
  # NOTE: particpant could have observations recorded in different observation files
  # this can create duplicate deficit by time_point; thus if deficit is recorded within same time_point but in different observation files
  # so (use max deficit) by time_point
  group_by(patid,setid) |>
     mutate(across(ends_with("bl|yr[1-5]"),~max(.x,na.rm = T))) |>
     ungroup() |>
    funique(cols=c("patid","setid")) 
    
```

### Polypharmacy eFI deficit

1. Create polypharmacy deficit from drug files
2. Define polypharmacy as ≥5 unique prescribed medications, using chapters 1–15 of the British National Formulary
define polypharmacy within one year intervals; CPRD study:https://www.cprd.com/approved-studies/redefining-polypharmacy-population-based-longitudinal-study-using-electronic

Approach:

1. Start with drugsssue files
2. Filter product codes in aurum product dictionary
3. Create study time points
4. Within each time interval, count n_distinct BNFChapters
5. If n>=5 plypharmacy==1 otherwise 0

```{r eFI_deficit_polypharmacy}

# Iterate the ff code on the drug issue files
map(drugissues,.progress = T,function(x) {
      collect(x)|>
      fselect(patid,issuedate,prodcodeid) |>
      fmutate(issuedate=as.Date(issuedate,"%d/%m/%Y")) |>
      join(collect(aurum_product_dictionary),on=c("prodcodeid"),how="inner") |> 
      fsubset(!is.na(BNFChapter)) |>
      join(select(collect(cohort_matched_2),patid,setid,indexdate,enddate),
       on=c("patid"),
       how="left",
       multiple = TRUE) |>
  # create polypharmacy deficit
    fmutate(time_point = case_when(issuedate > as.Date(-Inf) & issuedate <= indexdate-14 ~ "bl",
                                   issuedate > indexdate & issuedate <= indexdate+365 ~ "yr1",
                                   issuedate > indexdate+365 & issuedate <= indexdate+730 ~ "yr2",
                                   issuedate > indexdate+730 & issuedate <= indexdate+1095 ~ "yr3",
                                   issuedate > indexdate+1095 & issuedate <= indexdate+1460 ~ "yr4",
                                   issuedate > indexdate+1460 & issuedate <= indexdate+1825 ~ "yr5")) |>
     fsubset(!is.na(time_point)) |>
     group_by(patid,setid) |>
     fmutate(polypharmacy_n = n_distinct(BNFChapter)) |>
     funique(cols=c("patid","setid","time_point")) |>
     fselect(patid,setid,time_point,polypharmacy_n)}) 

# 
 polypharmacy_counts_list |>
  reduce(rowbind) |>
   # sum polypharmacy_n
   group_by(patid,setid,time_point) |>
   fmutate(polypharmacy_n = sum(polypharmacy_n,na.rm = T)) |>
   ungroup() |>
  funique(cols=c("patid","setid","time_point")) |>
  fmutate(polypharmacy = if_else(polypharmacy_n >= 5, 1, 0)) |>
  fmutate(polypharmacy_bl = if_else(polypharmacy == 1 & time_point == "bl",1,0)) |>
  fmutate(polypharmacy_yr1 = if_else(polypharmacy == 1 & time_point == "yr1",1,polypharmacy_bl)) |>
  fmutate(polypharmacy_yr2 = if_else(polypharmacy == 1 & time_point == "yr2",1,polypharmacy_yr1)) |>
  fmutate(polypharmacy_yr3 = if_else(polypharmacy == 1 & time_point == "yr3",1,polypharmacy_yr2)) |>
  fmutate(polypharmacy_yr4 = if_else(polypharmacy == 1 & time_point == "yr4",1,polypharmacy_yr3)) |>
  fmutate(polypharmacy_yr5 = if_else(polypharmacy == 1 & time_point == "yr5",1,polypharmacy_yr4)) |>
  funique(cols=c("patid","setid")) |>
  # individuals with no drug issue dates between study period or not match polyhpharmacy definition, polypharmacy == 0   
  fmutate(polypharmacy_not_missing = 1) |>
  join(fselect(collect(cohort_matched_2),patid,setid,indexdate,enddate),
       on=c("patid","setid"),
       how="full") |>
  fmutate(polypharmacy_bl = if_else(is.na(polypharmacy_not_missing),0,polypharmacy_bl),
          polypharmacy_yr1 = if_else(is.na(polypharmacy_not_missing),0,polypharmacy_yr1),
          polypharmacy_yr2 = if_else(is.na(polypharmacy_not_missing),0,polypharmacy_yr2),
          polypharmacy_yr3 = if_else(is.na(polypharmacy_not_missing),0,polypharmacy_yr3),
          polypharmacy_yr4 = if_else(is.na(polypharmacy_not_missing),0,polypharmacy_yr4),
          polypharmacy_yr5 = if_else(is.na(polypharmacy_not_missing),0,polypharmacy_yr5)) |>
  # censored individuals should have missing deficit
  fmutate(person_time_days = as.numeric(enddate - indexdate)) |>
  fmutate(polypharmacy_yr1 = if_else(person_time_days > 0,polypharmacy_yr1,NA_integer_),
          polypharmacy_yr2 = if_else(person_time_days > 365,polypharmacy_yr2,NA_integer_),
          polypharmacy_yr3 = if_else(person_time_days > 365*2,polypharmacy_yr3,NA_integer_),
          polypharmacy_yr4 = if_else(person_time_days > 365*3,polypharmacy_yr4,NA_integer_),
          polypharmacy_yr5 = if_else(person_time_days > 365*4,polypharmacy_yr5,NA_integer_)) |>
  fselect(patid,setid,polypharmacy_bl,polypharmacy_yr1,polypharmacy_yr2,
          polypharmacy_yr3,polypharmacy_yr4,polypharmacy_yr5) 
```

### Combine 35 deficits with polypharmacy deficit
```{r all_eFI_deficits}

# all_eFI_deficits
  eFI_35_deficits_data |>
  join(polypharmacy_data,
            on=c("patid","setid"),
       how="left")
```

### Create eFI 

Approach:

1. Sum deficits ==1 and divide by 36

```{r eFI_tidy}

# create eFI variable
  eFI_deficits_data|>
  mutate(eFI_bl = rowSums(pick(matches(glue("^{deficits_names}_bl$"))))/36) |> 
  mutate(eFI_yr1 = rowSums(pick(matches(glue("^{deficits_names}_yr1$"))))/36) |>
  mutate(eFI_yr2 = rowSums(pick(matches(glue("^{deficits_names}_yr2$"))))/36) |>
  mutate(eFI_yr3 = rowSums(pick(matches(glue("^{deficits_names}_yr3$"))))/36) |>
  mutate(eFI_yr4 = rowSums(pick(matches(glue("^{deficits_names}_yr4$"))))/36) |>
  mutate(eFI_yr5 = rowSums(pick(matches(glue("^{deficits_names}_yr5$"))))/36) |>
  # create eFI categories
  mutate(across(matches("^eFI_(bl|yr[1-5])$"),
                ~cut(.x,
                    breaks = c(0,0.12,0.24,0.36,Inf),
                    include.lowest = T,
                    labels = c("Fit: 0-0.12",
                               "Mild: >0.12-0.24",
                               "Moderate: >0.24-0.36",
                               "Severe: >0.36")),
                .names="{.col}_cat")) |>
  fselect(patid,setid,eFI_bl,eFI_bl_cat,
         eFI_yr1,eFI_yr1_cat,
         eFI_yr2,eFI_yr2_cat,
         eFI_yr3,eFI_yr3_cat,
         eFI_yr4,eFI_yr4_cat,
         eFI_yr5,eFI_yr5_cat,
         atrial_fibrillation_bl,cerebrovascular_disease_bl,
         heart_failure_bl,heart_valve_disease_bl,
         rheumatoid_diseases_bl,
         chronic_kidney_disease_bl,hypertension_bl,
         ischaemic_heart_disease_bl,urinary_system_disease_bl,
         osteoporosis_bl,parkinsonism_tremors_bl) 
```

## Frailty data 2: Create eFI deficits at baseline and overtime, thus as when new deficits are recorded at any time; for multi-state Cox regression

Approach:
1. Start with observation file
2. Join observation file with codelist on medcodeid,
3. If code matches, deficit = 1, otherwise deficit = 0
4. If patid censored, deficit = NA otherwise bullet 3
5. First define eFI at index date and update at each time a new (i.e, different from baseline) deficit is recorded, 
6. Will use the deficit at indexdate from that from the annual deficits;
so here is to just identify the new deficits and their dates
6. 36 deficits in total, 35 are are commordities/health etc defined with codelist and 
poypharmacy defined by counting number of prescribed medicines in BNF chapters 1-15 

### Polypharmacy eFI deficit mstate

1. Create polypharmacy deficit from drug files
2. Define polypharmacy as ≥5 prescribed medications, using chapters 1–15 of the British National Formulary
define polypharmacy within one year intervals; CPRD study:https://www.cprd.com/approved-studies/redefining-polypharmacy-population-based-longitudinal-study-using-electronic

Approach:
  
1. Start with drugsssue files
2. Filter product codes in aurum product dictionary
3. All needed is the first time polypharmacy was recorded
4. This will be the date the fifth drug was issued

```{r eFI_deficit_polypharmacy_mstate}

# Iterate the ff code on the drug issue files
new_polypharmacy_data <- map(drugissues,.progress = T,function(x) {
   collect(x)|>
    fselect(patid,issuedate,prodcodeid) |>
    fmutate(issuedate=as.Date(issuedate,"%d/%m/%Y")) |>
    join(collect(aurum_product_dictionary),on=c("prodcodeid"),how="inner") |> 
    fsubset(!is.na(BNFChapter)) |>
    join(select(collect(cohort_matched_2),patid,setid,indexdate,enddate),
         on=c("patid"),
         how="left",
         multiple = TRUE) |>
    fsubset(issuedate > as.Date(-Inf) & issuedate <= enddate) |>
    fselect(patid,setid,indexdate,issuedate,BNFChapter) |>
    rename(obsdate=issuedate) |> 
    distinct(patid,setid,BNFChapter,.keep_all = T)}) |> 
  reduce(rowbind,fill=T) |> 
  distinct(patid,setid,BNFChapter,.keep_all = T) |> 
  group_by(patid,setid) |>
  filter(n()>=5) |> 
  arrange(obsdate,.by_group = T) |>
  slice(5) |> 
  fselect(patid,setid,obsdate,indexdate) |>
  ungroup() |>
  arrange(patid,setid,obsdate) |>
  filter(obsdate > indexdate) |>
  fselect(patid,setid,obsdate) |>
  to_duckdb(con=my_con_integer64)
```

### 36 eFI deficits for multistate analyses

```{r 36_eFI_deficits_mstate}

# file paths to codelist
# use list.files() function to get the list of file paths
eFI_deficits_codelists_file_paths

# a vector of the names of deficits
deficits_names <- map_chr(eFI_deficits_codelists_file_paths, ~str_extract(.x,pattern = "(.*/)(.*)( .* codelist.csv$)",group = 2))

# a named list of the codelists
deficits_code_list <- setNames(map(eFI_deficits_codelists_file_paths, ~unique(pull(collect(open_csv_dataset(.x)),medcodeid))),
                               deficits_names)


# all codelits combined: vector
deficits_code_df <- map(deficits_code_list,~as.data.frame(.x)) |> reduce(rowbind) |>
  rename_with(~"medcodeid")

# define deficit based on codelist
# iterate the ff code on the observation files and combine result into a data set of the 35 deficits

map(observations,.progress = TRUE,function(x) {
  
    collect(x) |> 
    fselect(patid,obsdate,medcodeid) |>
    funique(cols=c("patid","obsdate","medcodeid")) |>
    fmutate(obsdate = as.Date(obsdate,format="%d/%m/%Y")) |>
    fsubset(medcodeid %in% deficits_code_df$medcodeid) |>
    fmutate(deficit_name=case_when(medcodeid %in% deficits_code_list[[1]] ~ deficits_names[[1]],
                                   medcodeid %in% deficits_code_list[[2]] ~ deficits_names[[2]],
                                   medcodeid %in% deficits_code_list[[3]] ~ deficits_names[[3]],
                                   medcodeid %in% deficits_code_list[[4]] ~ deficits_names[[4]],
                                   medcodeid %in% deficits_code_list[[5]] ~ deficits_names[[5]],
                                   medcodeid %in% deficits_code_list[[6]] ~ deficits_names[[6]],
                                   medcodeid %in% deficits_code_list[[7]] ~ deficits_names[[7]],
                                   medcodeid %in% deficits_code_list[[8]] ~ deficits_names[[8]],
                                   medcodeid %in% deficits_code_list[[9]] ~ deficits_names[[9]],
                                   medcodeid %in% deficits_code_list[[10]] ~ deficits_names[[10]],
                                   medcodeid %in% deficits_code_list[[11]] ~ deficits_names[[11]],
                                   medcodeid %in% deficits_code_list[[12]] ~ deficits_names[[12]],
                                   medcodeid %in% deficits_code_list[[13]] ~ deficits_names[[13]],
                                   medcodeid %in% deficits_code_list[[14]] ~ deficits_names[[14]],
                                   medcodeid %in% deficits_code_list[[15]] ~ deficits_names[[15]],
                                   medcodeid %in% deficits_code_list[[16]] ~ deficits_names[[16]],
                                   medcodeid %in% deficits_code_list[[17]] ~ deficits_names[[17]],
                                   medcodeid %in% deficits_code_list[[18]] ~ deficits_names[[18]],
                                   medcodeid %in% deficits_code_list[[19]] ~ deficits_names[[19]],
                                   medcodeid %in% deficits_code_list[[20]] ~ deficits_names[[20]],
                                   medcodeid %in% deficits_code_list[[21]] ~ deficits_names[[21]],
                                   medcodeid %in% deficits_code_list[[22]] ~ deficits_names[[22]],
                                   medcodeid %in% deficits_code_list[[23]] ~ deficits_names[[23]],
                                   medcodeid %in% deficits_code_list[[24]] ~ deficits_names[[24]],
                                   medcodeid %in% deficits_code_list[[25]] ~ deficits_names[[25]],
                                   medcodeid %in% deficits_code_list[[26]] ~ deficits_names[[26]],
                                   medcodeid %in% deficits_code_list[[27]] ~ deficits_names[[27]],
                                   medcodeid %in% deficits_code_list[[28]] ~ deficits_names[[28]],
                                   medcodeid %in% deficits_code_list[[29]] ~ deficits_names[[29]],
                                   medcodeid %in% deficits_code_list[[30]] ~ deficits_names[[30]],
                                   medcodeid %in% deficits_code_list[[31]] ~ deficits_names[[31]],
                                   medcodeid %in% deficits_code_list[[32]] ~ deficits_names[[32]],
                                   medcodeid %in% deficits_code_list[[33]] ~ deficits_names[[33]],
                                   medcodeid %in% deficits_code_list[[34]] ~ deficits_names[[34]],
                                   medcodeid %in% deficits_code_list[[35]] ~ deficits_names[[35]])) |>
    fselect(patid,obsdate,deficit_name) |>
    join(fselect(collect(cohort_matched_2),patid,setid,indexdate,enddate),
         on = "patid",
         how ="left",
         multiple = TRUE) |> 
    fsubset(obsdate > as.Date(-Inf) & obsdate <= enddate) |> 
    arrange(patid,setid,obsdate) |>
    funique(cols=c("patid","setid","deficit_name")) |> 
    filter(obsdate > indexdate) |>
    fselect(patid,setid,obsdate) |>
    rbind(collect(new_polypharmacy_data)) |>
    arrange(patid,setid,obsdate) |>
    group_by(patid,setid,obsdate) |>
    mutate(n_new_deficits_onday=n()) |> 
    ungroup() |>
    distinct(patid,setid,obsdate,.keep_all = T) }) |> 
  reduce(rowbind,fill=T) |> 
  distinct(patid,setid,obsdate,.keep_all = T) |> 
  arrange(patid,setid,obsdate) 
```

## Now create frailty states over time

1. Recalculate eFI anytime a new deficit is recorded; updating the previous eFI (noting the number of new deficits occurring exactly in same obsdate);
Thus, group by c(patid,setid,obsdate) and then recalculate eFI by group after indexdate
2. Based on that create a new status only when new eFI category is different from the previous

```{r eFI_multi_states}

# extract baseline eFI score
eFI_score_bl <- eFI_data |>
   select(patid,setid,eFI_bl) |>
   left_join(transmute(collect(cohort_matched_2),patid,setid,eFI_bl_date =indexdate))

# Create eFI mstate: note: any new one deficit increases eFI_bl by 1/36 = 0.02777778
eFI_mstate_data <- new_eFI_36_deficits_data |>
  group_by(patid,setid) |>
  arrange(patid,setid,obsdate,.by_group = T) |>
  ungroup() |>
  left_join(eFI_score_bl) |>
  group_by(patid,setid) |>
  arrange(obsdate,.by_group = T) |>
  mutate(eFI_mstate = eFI_bl+(cumsum(n_new_deficits_onday)*0.02777778)) |>
  # create categories
  ungroup() |>
  select(patid,setid,obsdate,eFI_mstate) |>
  bind_rows(transmute(eFI_score_bl,patid,setid,obsdate=eFI_bl_date,eFI_mstate =eFI_bl)) |> 
  mutate(eFI_mstate_cat =  cut(eFI_mstate,
                    breaks = c(0,0.12,0.24,0.36,Inf),
                    include.lowest = T,
                    labels = c("Fit: 0-0.12",
                               "Mild: >0.12-0.24",
                               "Moderate: >0.24-0.36",
                               "Severe: >0.36"))) |>
  fselect(patid,setid,eFI_mstate_cat,obsdate) |>
  #keep the first of each eFI cat
  arrange(patid,setid,obsdate) |>
  distinct(patid,setid,eFI_mstate_cat,.keep_all = T) |>
  # for each observation, when did each of the unique eFI_cat occur and date
  mutate(eFI_fit = case_when(eFI_mstate_cat == "Fit: 0-0.12" ~ 1),
         eFI_mild = case_when(eFI_mstate_cat == "Mild: >0.12-0.24" ~ 1),
         eFI_moderate = case_when(eFI_mstate_cat == "Moderate: >0.24-0.36" ~ 1),
         eFI_severe = case_when(eFI_mstate_cat == "Severe: >0.36" ~ 1)) |>
  mutate(eFI_fit_date = case_when(eFI_mstate_cat == "Fit: 0-0.12" ~ obsdate),
         eFI_mild_date = case_when(eFI_mstate_cat == "Mild: >0.12-0.24" ~ obsdate),
         eFI_moderate_date = case_when(eFI_mstate_cat == "Moderate: >0.24-0.36" ~ obsdate),
         eFI_severe_date = case_when(eFI_mstate_cat == "Severe: >0.36" ~ obsdate)) |>
  group_by(patid,setid) |>
  fill(matches("eFI_(fit|mild|moderate|severe)(_date)?"),.direction="downup") |>
  ungroup() |>
  select(patid,setid,matches("eFI_(fit|mild|moderate|severe)(_date)?")) |>
  distinct(patid,setid,.keep_all = T) |>
  left_join(collect(transmute(cohort_matched_2,patid,setid,indexdate,enddate,eFI_death_date=cprd_ddate))) |>
  mutate(eFI_death = case_when(eFI_death_date <= enddate ~ 1,.default = 0)) |>
  mutate(eFI_death_date = case_when(eFI_death_date <= enddate ~ eFI_death_date)) |>
  mutate(eFI_fit_days = NA_real_) |> 
  mutate(eFI_mild_days = as.numeric(pmin(eFI_mild_date,enddate,na.rm = T)-indexdate)) |> 
  mutate(eFI_moderate_days = as.numeric(pmin(eFI_moderate_date,enddate,na.rm = T)-indexdate)) |>
  mutate(eFI_severe_days =  as.numeric(pmin(eFI_severe_date,enddate,na.rm = T)-indexdate)) |>
  mutate(eFI_death_days = as.numeric(pmin(eFI_death_date,enddate,na.rm = T)-indexdate)) |>
  # IF DEATH AND eFI ON SAME DATE, DEATH SHOULD TAKE PRECEDENCE
  mutate(across(matches("^eFI_(fit|mild|moderate|severe)$"),
         ~if_else(get(glue("{cur_column()}_date"))==enddate,0,.x))) |>
  mutate(across(matches("^eFI_(fit|mild|moderate|severe)_date$"),
         ~if_else(.x==enddate,NA_Date_,.x))) |>
  # Convert time to years
  mutate(across(matches("_days$"),~.x/365.25,.names="{.col}_years")) |>
  rename_with(~str_replace(.x,"days_years","years")) |>
  relocate(patid,setid,indexdate,enddate) |>
  mutate(across(matches("^eFI_(fit|mild|moderate|severe)$"),
                ~if_else(is.na(.x),0,.x))) |>
  relocate(eFI_death,.after = eFI_severe) |>
  relocate(eFI_death_date,.after = eFI_severe_date) |>
  to_duckdb(con = my_con_integer64)

```


# Study flow chart
```{r flow_chart}

# Study flow chart

# All individuals in denominator file aged > 65 years before study ends in 2019-12-31

all_aged_65 <- denominator |> # denominator dataset
  select(patid, yob) |>
  collect() |>
  mutate(birthday_65th=as.Date(paste0(yob+65, "/06/01"),format = "%Y/%m/%d")) |> 
  filter(birthday_65th<=as.Date("2019-12-31","%Y-%m-%d")) |>
  nrow()

# Eligible for inclusion

nrow(collect(cohort_eligible))

# Not eligible for inclusion

all_aged_65 - nrow(collect(cohort_eligible))

# Matched
nrow(collect(cohort_matched_2))
tabyl(collect(cohort_matched_2),exposed)

# Not matched
nrow(collect(cohort_eligible)) - nrow(collect(cohort_matched_2))

# XXX had missing frailty index (eFI) from 1-5 years follow-up (were censored at cohort entry)
eFI_data |>
  filter(!is.na(eFI_bl),if_all(matches("^eFI_yr[1-5]$"),~is.na(.x))) |> 
  dim()

# XXX had missing frailty index (eFI) from 1-5 years follow-up (were censored at cohort entry)
# this does not apply as if not censored, eFI is either yes or no

# Final cohort
collect(cohort_matched_2)|>
  left_join(eFI_data.rds, join_by(patid,setid)) |>
  filter(if_any(matches("^eFI_yr[1-5]$"),~!is.na(.x))) |> 
  count(exposed)

# Infection site and type
collect(cohort_matched_2)|>
  left_join(eFI_data.rds, join_by(patid,setid)) |>
  filter(if_any(matches("^eFI_yr[1-5]$"),~!is.na(.x)),exposed==1) |> 
  left_join(infection_data , join_by(patid,setid)) |>
  count(inf_type_of_match_grp )
#
collect(cohort_matched_2)|>
  left_join(eFI_data , join_by(patid,setid)) |>
  filter(if_any(matches("^eFI_yr[1-5]$"),~!is.na(.x)),exposed==1) |> 
  left_join(infection_data, join_by(patid,setid)) |>
  count(inf_site_of_match_grp)
```

